---
title: "plotting.qmd"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Intro
Disclaimer: this code is a mess, doesn't follow DRY principles, and I wrote it in a hurry. But it works!
If you're a prospective employer, please don't judge me too much and know that I'm able to write clean code when I need to :P

# Libraries
```{r}

# working_env  <- "genomedk"
working_env  <- "local"
date  <- "29_04_2025"

final_folder  <- "/Users/lmz306/Library/CloudStorage/OneDrive-UniversityofCopenhagen/igpv/eir_psy/results/2024-05-28-Plots_paper/paper_final_29_04_2025"

library(tidyverse)
library(ggpattern)
library(glue)
library(rstatix)
library(ggbreak)
library(ggrepel)
library(tidymodels)
library(readxl)
library(tidytext) # for scale_x_reordered()
library(patchwork)
library(openxlsx)
library(plotly)
library(gt) 
library(gtsummary)
library(GGally)
library(ggpmisc)
library(glue)
library(ggpubr)
library(ggplot2)
library(knitr)
library(kableExtra)
# library(gt)
library(insight)
library(ggpattern)
library(rlang)

figs_path  <- "/Users/lmz306/Library/CloudStorage/OneDrive-UniversityofCopenhagen/igpv/eir_psy/results/2024-05-28-Plots_paper/figs_paper_28_04_25"




#-
my_theme <- theme_bw() + 
theme(
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  plot.title = element_text(hjust = 0.5, color = "black"),
  text=element_text(size=30, color = "black"),
  panel.border = element_rect(linewidth = 2, color = "black"),
  strip.background = element_blank(),
  axis.ticks.x = element_blank(),  # Remove x ticks
  axis.text.y.right = element_blank(),
  axis.line.y.right = element_blank(),
  axis.ticks.y.right = element_blank(),
  # center x label under the tick
    # axis.text.x = element_text(hjust = 0.5),
  # axis.ticks.x = element_blank(),
  # change all text color to black
  axis.text = element_text(color = "black"),
  axis.text.x = element_text(hjust = 0.5),
  legend.key = element_rect(linewidth = unit(1, "cm")),  # Set fixed size for legend keys
  legend.key.size = unit(1, "cm"),                  # Make legend keys square
  legend.key.height = unit(1, "cm"),                # Ensure height matches width
  legend.key.width = unit(1, "cm")                  # Ensure width matches height
  # axis.text.x = element_text(color = "black"),
)
# Do not set it to not mess with ggbreak
# theme_set(my_theme)
suffix_delta_eval  <- " in iPSYCH2"
suffix_delta_model  <- " with DL"

current_dir  <- file.path(".")

if (working_env == "local"){
  message("Working environment: local")
  setwd("/Users/lmz306/Library/CloudStorage/OneDrive-UniversityofCopenhagen/igpv/eir_psy/results/2024-05-28-Plots_paper")
  source("/Users/lmz306/Library/CloudStorage/OneDrive-UniversityofCopenhagen/igpv/scripts/functions.R")
  source("plotting_functions.R")
  remove_ADHD  <- FALSE
} else if (working_env == "genomedk"){
  message("Working environment: genomedk")
  setwd("/faststorage/jail/project/igpv/eirpsy/results/2024-05-28-Plots_paper")
  source("/faststorage/jail/project/igpv/scripts/functions.R")
  source("plotting_functions.R")
  remove_ADHD  <- FALSE
}

if(!dir.exists("plots_paper_final")){
  dir.create("plots_paper_final")
}

if(!dir.exists(glue("plots_paper_final/data_{date}"))){
  dir.create(glue("plots_paper_final/data_{date}"))
}
if(!dir.exists(glue("plots_paper_final/figs_{date}"))){
  dir.create(glue("plots_paper_final/figs_{date}"))
}

hypers_changed  <- c(
"kernel_width_input_gln_config", "channel_exp_base_input_gln_config", "rb_do_input_gln_config", "na_augment_perc_input_gln_config")
hypers_changed_short  <- str_replace(hypers_changed, "_input_gln_config", "")

add_significance_leo <- function(data, sig_col_name, p_value_col_name) {
  data %>%
    mutate(!!sig_col_name := case_when(
      !!sym(p_value_col_name) < 0.001 ~ "***",
      !!sym(p_value_col_name) < 0.01 ~ "**",
      !!sym(p_value_col_name) < 0.05 ~ "*",
      TRUE ~ "ns"
    ))
}



add_all_significance_columns <- function(data) {
  data %>%
    add_significance_leo(sig_col_name = "sig_nom", p_value_col_name = "p_value") %>% 
    add_significance_leo(sig_col_name = "sig_bonf_diag", p_value_col_name = "p_value_bonf_diag") %>% 
    add_significance_leo(sig_col_name = "sig_fdr_diag", p_value_col_name = "p_value_fdr_diag") %>% 
    add_significance_leo(sig_col_name = "sig_bonf_all", p_value_col_name = "p_value_bonf_all") %>% 
    add_significance_leo(sig_col_name = "sig_fdr_all", p_value_col_name = "p_value_fdr_all")
}
filter_out_ADHD  <- function(df){
  if (remove_ADHD) {
    df <- df %>% filter(Diagnosis != "ADHD") %>% 
    mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))
  }
  return(df %>% 
    mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))
  )
}


metrics_r2 <- c("avg_r2", "std_r2", "lower_r2", "upper_r2")
metrics_auroc <- c("avg_roc_auc", "std_roc_auc", "lower_roc_auc", "upper_roc_auc")

metrics_auc  <- c("avg_auc", "sd_auc", "lower_auc", "upper_auc")
count_cols  <- function(df){
  # if there's column called Run and Model_type, include them in the count
  if ("Run" %in% colnames(df) & "Model_type" %in% colnames(df)){
    df %>% 
    count(Model, Features, Run, Model_type, Interaction, CV, Split_trained, Split_eval) %>% print(n=Inf)
  } else if ("Run" %in% colnames(df)){
    df %>% 
    count(Model, Features, Run, Interaction, CV, Split_trained, Split_eval) %>% print(n=Inf)
  } else if ("Model_type" %in% colnames(df)){
    df %>% 
    count(Model, Features, Model_type, Interaction, CV, Split_trained, Split_eval) %>% print(n=Inf)
  } else {
    df %>% 
    count(Model, Features, Interaction, CV, Split_trained, Split_eval) %>% print(n=Inf)
  }
}

  apply_correct_multiple <- function(df, correct_multiple = TRUE){
    if (correct_multiple == TRUE){
      # Do your per-diagnosis p-value correction
      df  <- df %>% group_by(Diagnosis) %>%
      mutate(
      p_value_bonf_diag = p.adjust(p_value, method = "bonferroni"),
      p_value_fdr_diag  = p.adjust(p_value, method = "fdr")
    ) %>%
    ungroup() %>%
    # Do your global p-value correction
    mutate(
      p_value_bonf_all = p.adjust(p_value, method = "bonferroni"),
      p_value_fdr_all  = p.adjust(p_value, method = "fdr")
    )
  } else {
    df <- df
  }
  return(df)
  }

#-
compute_pvalues_comparisons <- function(to_compare, data, columns_to_group, runs_to_compare, filter_column, save_plots = TRUE, metric = "auc", folder_name = folder_name, correct_multiple = TRUE) {
  require(dplyr)
  require(tidyr)
  require(purrr)
  require(glue)

  if (metric == "auc"){
    metric_col <- "auc_replicates"
  } else if (metric == "r2"){
    metric_col <- "r2_replicates"
  }

  # message(glue("columns_to_group: {columns_to_group}"))
  columns_to_group_list  <- list(columns_to_group)
  # message(glue("columns_to_group_list: {columns_to_group_list}"))

  data %>%
    # Only keep rows whose `filter_column` value is in any run pair
    group_by(across(all_of(columns_to_group))) %>% 
    nest() %>% 

    mutate(comparisons = pmap(list(data, !!!syms(columns_to_group)), function(group_data, ...) {
      
      map_dfr(runs_to_compare, function(pair) {
        run1_name <- pair[1]
        run2_name <- pair[2]

        # print(run1_name)
        # print(run2_name)
        
        # Filter the group_data based on the chosen filter_column
        run1_data <- group_data %>% filter(!!sym(filter_column) == run1_name)
        run2_data <- group_data %>% filter(!!sym(filter_column) == run2_name)

        # print(run1_data)
        # print(run2_data)

        if (nrow(run1_data) == 0 || nrow(run2_data) == 0) {
          message(glue("Skipping {run1_name} vs {run2_name}: missing data"))
          message("run1_data:")
          print(run1_data)
          message("run2_data:")
          print(run2_data)
          return(data.frame(Model1 = run1_name, Model2 = run2_name))
        }
        
        run1_reps <- list(replicates = unlist(run1_data[[metric_col]]))
        run2_reps <- list(replicates = unlist(run2_data[[metric_col]]))
        
        if (length(run1_reps$replicates) == 0 || length(run2_reps$replicates) == 0) {
          message(glue("Skipping {run1_name} vs {run2_name}: empty replicates"))
          message("run1_reps:")
          print(run1_reps)
          message("run2_reps:")
          print(run2_reps)
          return(data.frame(Model1 = run1_name, Model2 = run2_name))
        }

        # message(glue("CV = {CV}"))
        # message(glue("Split_trained = {Split_trained}"))
        # message(glue("Split_eval = {Split_eval}"))
        # message(glue("Diagnosis = {Diagnosis}"))
        
        # message("printing group_data")
        # print(group_data)

        # if "Integration" is not in the columns_to_group, assign it to NULL
        CV  <- if ("CV" %in% columns_to_group) CV else NULL
        Integration  <- if ("Integration" %in% columns_to_group) Integration else NULL
        Features     <- if ("Features" %in% columns_to_group) Features else NULL
        Model_type   <- if ("Model_type" %in% columns_to_group) Model_type else NULL
        Split_trained <- if ("Split_trained" %in% columns_to_group) Split_trained else NULL
        Split_eval <- if ("Split_eval" %in% columns_to_group) Split_eval else NULL
        
        comparison <- compare_metric_distributions(
          to_compare = to_compare,
          metric_boot1  = run1_reps,
          metric_boot2  = run2_reps,
          metric = metric,
          run1_name  = run1_name,
          run2_name  = run2_name,
          Diagnosis  = Diagnosis,
          CV         = CV,
          Features   = Features,
          Integration = Integration,
          Model_type = Model_type,
          Split_trained = Split_trained,
          Split_eval = Split_eval,
          save_plots = save_plots,
          folder_name = folder_name
        )
        
        # Extract stats
        comp_stats <- comparison$comparative_statistics
        comp_stats$Model1 <- run1_name
        comp_stats$Model2 <- run2_name
        
        comp_stats %>%
          as_tibble() %>%
          mutate(Value = as.numeric(Value)) %>%
          pivot_wider(names_from = Metric, values_from = Value)
      })
      
    })) %>% 
    select(-data) %>%
    unnest(comparisons) %>%
    ungroup() %>% 
    apply_correct_multiple(correct_multiple = correct_multiple)
}

#-
compare_metric_distributions <- function(to_compare, metric = "auc", metric_boot1, metric_boot2, run1_name = "Model_1", run2_name = "Model_2", folder_name = folder_name, Diagnosis = NULL, Integration = NULL, Model_type = NULL, Features = NULL, Split_trained = NULL, Split_eval = NULL, CV = NULL, bins = 100, save_plots = TRUE){

  library(ggplot2)
  library(tidyr)
  library(dplyr)

  # create folder if it doesn't exist
  if(!dir.exists(folder_name)){dir.create(folder_name, showWarnings = FALSE)}
  
  metric_col <- if (metric == "auc") "AUC" else "R2"
  # print("the function is running...")
  # Prepare data for plotting
  df <- data.frame(
    metric_col = c(metric_boot1$replicates, metric_boot2$replicates),
    Model = factor(c(rep(run1_name, length(metric_boot1$replicates)),
                    rep(run2_name, length(metric_boot2$replicates))), levels = c(run1_name, run2_name))
  ) %>% as_tibble()
  
    # Create histogram data to get max height for label positioning
  hist_data <- ggplot_build(
    ggplot(df, aes(x = metric_col)) +
    geom_histogram(bins = bins)
  )$data[[1]]
  max_height <- max(hist_data$count) * 1.1  # 10% above max height
  
  # Calculate y-position for horizontal lines (middle of the plot)
  mid_height <- max_height / 2

  # Calculate comprehensive statistics
  stats <- df %>%
    group_by(Model) %>%
    summarise(
      mean_metric = mean(metric_col, na.rm = TRUE),
      sd_metric = sd(metric_col, na.rm = TRUE),
      median_metric = median(metric_col, na.rm = TRUE),
      ci_lower = quantile(metric_col, 0.025, na.rm = TRUE),
      ci_upper = quantile(metric_col, 0.975, na.rm = TRUE),
      n_samples = n()
    ) %>% 
  mutate(label_x = ifelse(Model == levels(Model)[1], mean_metric - 0.02, mean_metric + 0.02)) %>% 
  mutate(y_offset = ifelse(Model == levels(Model)[1], mid_height * 1.02, mid_height * 0.98))

  
  # Calculate direct p-value
  direct_p_value <- mean(
    map_dbl(metric_boot2$replicates, ~ mean(metric_boot1$replicates >= .x))
  )

  mean_diff  <- stats$mean_metric[2] - stats$mean_metric[1]
  # Here every bootstrap from run1 is compared to every bootstrap from run2. So it counts how many times the AUC from run1 is greater than the AUC from run2 and then takes the mean of that, effectively calculating the p-value of the Null hypothesis that run1 is better than run2.


  
  # Calculate effect size (Cohen's d)
  # pooled_sd <- sqrt((stats$sd_metric[1]^2 + stats$sd_metric[2]^2) / 2)
  # cohens_d <- (stats$mean_metric[2] - stats$mean_metric[1]) / pooled_sd
  
  # print(df)
  diagnosis_text <- if (!is.null(Diagnosis)) Diagnosis else ''
  integration_text <- if (!is.null(Integration)) paste0(Integration, ' integration') else ''
  cv_text <- if (!is.null(CV)) paste0('CV = ', CV) else ''
  split_trained_text <- if (!is.null(Split_trained)) paste0('Split trained = ', Split_trained) else ''
  split_eval_text <- if (!is.null(Split_eval)) paste0('Split eval = ', Split_eval) else ''
  run2_name_text <- if (!is.null(run2_name)) paste0(run2_name, ' mean ', metric_col, ': ', format(stats$mean_metric[2], digits = 3)) else ''
  run1_name_text <- if (!is.null(run1_name)) paste0(run1_name, ' mean ', metric_col, ': ', format(stats$mean_metric[1], digits = 3)) else ''
  model_type_text <- if (!is.null(Model_type)) paste0('Model type = ', Model_type) else ''
  features_text <- if (!is.null(Features)) Features else ''

  subtitle_text <- glue("Test for {diagnosis_text} {model_type_text} {features_text} {integration_text} {cv_text} 
          {split_trained_text} {split_eval_text} 
          {run2_name_text} 
          {run1_name_text} 
          Mean difference: {format(mean_diff, digits = 3)} 
          Empirical p-value: {format(direct_p_value, digits = 3, scientific = TRUE)} ({format(direct_p_value, digits = 3, scientific = FALSE)})")

  
  integration_filename  <- if (!is.null(Integration)) paste0(Integration, '_') else ''
  CV_filename  <- if (!is.null(CV)) paste0(CV, '_') else ''
  Split_trained_filename  <- if (!is.null(Split_trained)) paste0(Split_trained, '_') else ''
  # Split_eval_filename  <- if (!is.null(Split_eval)) paste0(Split_eval) else ''
  Split_eval_filename  <- if (!is.null(Split_eval) && Split_eval != "") paste0(Split_eval, '_') else ''
  Model_type_filename  <- if (!is.null(Model_type) && Model_type != "") paste0(Model_type, '_') else ''

  features_filename  <- if (!is.null(Features)) paste0(Features, '_') else ''

  # message(glue("subtitle_text: {subtitle_text}"))
  # message(glue("run2_name: {run2_name}"))
  # message(glue("run1_name: {run1_name}"))
  # message(glue("CV_filename: {CV_filename}"))
  # message(glue("Split_trained_filename: {Split_trained_filename}"))
  # message(glue("Split_eval_filename: {Split_eval_filename}"))
  # message(glue("integration_filename: {integration_filename}"))
  # message(glue("Diagnosis: {Diagnosis}"))

  filename  <- glue("{metric_col}_comparison_{Diagnosis}_{Model_type_filename}{features_filename}{integration_filename}{run2_name}_vs_{run1_name}_{CV_filename}{Split_trained_filename}{Split_eval_filename}.png")
  # strip the filename from the last "_" before the extension if there's any
  filename  <- str_replace(filename, "_([^_]*)$", "\\1")
  # message(glue("filename: {filename}"))
    
  # Create plot with histograms
  # pdf(NULL)
  p <- ggplot(df, aes(x = metric_col, fill = Model)) +
    geom_histogram(alpha = 0.5, position = "identity", bins = bins) +
    # Add vertical lines for means
    geom_vline(data = stats, aes(xintercept = mean_metric, color = Model),
               linetype = "dashed", linewidth = 1) +
    # Add horizontal lines with mean and CI at mid-height
    geom_segment(data = stats,
                aes(x = ci_lower, xend = ci_upper,
                    y = y_offset, yend = y_offset,
                    color = Model),
                linewidth = 2, alpha = 0.5) +
    # Add point for mean on horizontal line
    geom_point(data = stats,
               aes(x = mean_metric, y = y_offset, color = Model),
               size = 3) +
    # Add mean labels with fixed y position
    # geom_label(data = stats,
    #           aes(x = mean_auc + c(-0.04, +0.04), y = mid_height + 3,
    #               label = sprintf("Mean: %.3f", mean_auc), fill = Model),
    #           color = "white", vjust = 0, hjust = -0.1, label.size = 0.25, linewidth = 4) +
    # geom_text_repel(data = stats,
    #             aes(x = mean_auc, y = mid_height + 3,
    #                 label = sprintf("Mean: %.3f", mean_auc), fill = Model),
    #             color = "white", box.padding = 0.35, point.padding = 0.5,
    #             segment.color = "grey50", linewidth = 4, min.segment.length = 0.1) + 
    geom_label_repel(
      data = stats,
      aes(
        x = mean_metric, 
        # y = mid_height + max_height * 0.05, # Position label slightly above mid-height
        y = y_offset,
        label = sprintf("Mean: %.3f", mean_metric),
        fill = Model # Use fill to create colored boxes for each model
      ),
      color = "white", # Text color
      box.padding = 0.35,
      point.padding = 0.5,
      segment.color = "grey50",
      size = 4,
      min.segment.length = 0.1,
      label.size = 0.25, # Border thickness of the box
      label.r = unit(0.2, "lines") # Rounded corners for the box
    ) + 
    scale_fill_manual(values = c(cbp1[1], cbp1[2]), guide = "none") +
    scale_color_manual(values = c(cbp1[1], cbp1[2])) +
    labs(x = glue("Mean {metric_col}"),
         y = "Count",
         title = glue("{Diagnosis} {model_type_text} {features_text} Testing H1:{run2_name} > {run1_name} "),
         subtitle = subtitle_text) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    # Ensure labels are visible by extending y axis
    coord_cartesian(clip = "off") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.2)))



#   filename  <- glue("{metric_col}_comparison_{Diagnosis}_{run2_name}_vs_{run1_name}_{CV}_{Split_trained}_{Split_eval}.pdf")

  if (save_plots == TRUE){
    message("Saving plot to ", file.path(folder_name, filename))
    # message("Saving plot to ", filename)

    # Save plot to file in folder_name
    # suppressMessages(
    # ggsave(file.path(folder_name, filename), p, device = "png", width = 11, height = 8.5, units = "in")
    ggsave(file.path(folder_name, filename), p + theme(plot.background = element_rect(fill = "white", color = NA)), device = "png", width = 11, height = 8.5, units = "in")

    # )
    # dev.off()
  }

  # Create summary statistics table
  summary_stats <- data.frame(
    Metric = c(
      "Mean {metric_col}",
      "Median {metric_col}",
      "Standard Deviation",
      "95% CI Lower",
      "95% CI Upper",
      "Number of Bootstraps"
    ),
    Model1 = format(c(
      stats$mean_metric[1],
      stats$median_metric[1],
      stats$sd_metric[1],
      stats$ci_lower[1],
      stats$ci_upper[1],
      stats$n_samples[1]
    ), digits = 4),
    Model2 = format(c(
      stats$mean_metric[2],
      stats$median_metric[2],
      stats$sd_metric[2],
      stats$ci_lower[2],
      stats$ci_upper[2],
      stats$n_samples[2]
    ), digits = 4)
  )
  
  names(summary_stats)[2:3] <- c(run1_name, run2_name)
  
  # Additional comparative statistics
  comparative_stats <- tibble(
    Metric = c(
      "Mean_diff",
      "p_value"
    ),
    Value = format(c(
      stats$mean_metric[2] - stats$mean_metric[1],
      direct_p_value
    ), digits = 4)
  ) %>% 
      mutate(H0 = "Model 2 <= Model 1",
      H1 = "Model 2 > Model 1")
  
  # Return all results
  return(invisible(list(
    plot = p,
    summary_statistics = summary_stats,
    comparative_statistics = comparative_stats,
    raw_stats = stats
  )))
}




set_comparison_parameters <- function(to_compare) {
  # Here the convention is that the second model to compare is tested to see if it is better than the first model, i.e. Model2 > Model1
  if (to_compare == "GLN_vs_bigstatsr") {
    columns_to_group  <- c("Diagnosis", "CV", "Split_trained", "Split_eval")
    runs_to_compare  <- list(c("bigstatsr_on_geno", "GLN_on_geno"))
    runs_to_compare_list  <- c("bigstatsr_on_geno", "GLN_on_geno")
  } else if (to_compare == "different_features_same_eval") {
    columns_to_group  <- c("Diagnosis", "Model", "Integration", "Model_type", "CV", "Split_trained", "Split_eval")
    runs_to_compare <- list(
      c("GLN", "GLN+bigstatsr"),
      c("PGS", "PGS+FGRS"),
      c("FGRS", "PGS+FGRS"),
      c("PGS+FGRS", "GLN+PGS+FGRS"),
      c("GLN+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS")
      # c("PGS+FGRS", "bigstatsr+PGS+FGRS"),
      # c("bigstatsr+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS"),
    )
    # runs_to_compare_list  <- c("PGS", "PGS+FGRS", "GLN+PGS+FGRS", "bigstatsr+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS")
    runs_to_compare_list  <- unlist(runs_to_compare)
  } else if (to_compare == "DL_vs_Linear") {
    columns_to_group  <- c("Diagnosis", "Features", "CV", "Split_trained", "Split_eval")
    runs_to_compare <- list(
      c("Linear", "DL")
    )
    runs_to_compare_list  <- c("Linear", "DL")
  } else if (to_compare == "same_feature_same_model_different_eval") {
    columns_to_group  <- c("Diagnosis", "Features", "CV", "Split_trained", "Model_type")
    runs_to_compare <- list(
      c("iPSYCH2", "test_iPSYCH1")
    )
    runs_to_compare_list  <- c("iPSYCH2", "test_iPSYCH1")
  }
  return(list(columns_to_group = columns_to_group, runs_to_compare = runs_to_compare, runs_to_compare_list = runs_to_compare_list))
}


#-
process_statistical_tests <- function(
    metrics_df,
    emp_pvalues_df,
    metric_name,
    x_col,
    fill_col,
    columns_to_group,
    nested_models = FALSE,
    step_increase = 0.12,
    keep_group_model_order = TRUE,
    print = FALSE,
    runs_to_compare = NULL,
    fun_to_apply = "max"
) {
    # Construct column names based on metric
    y_col <- paste0("avg_", metric_name)
    lower_col <- paste0("lower_", metric_name)
    upper_col <- paste0("upper_", metric_name)
    add_combination  <- function(df){
      if (nested_models == FALSE){
        df %>% 
          mutate(combination = paste(!!!syms(columns_to_group), sep = "_"))
      } else {
        df %>% 
          mutate(combination = paste(Diagnosis, group1, group2, sep = "_"))
      }
    }

    rename_groups  <- function(df){
      if (keep_group_model_order == TRUE){
        df %>% 
          rename(group1 = Model1, group2 = Model2)
      } else {
        df %>% 
          rename(group1 = Model2, group2 = Model1)
      }
    }

    metrics_df  <- metrics_df %>% 
        mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))
    
    # Create formula for t-test
    my_formula <- as.formula(paste(y_col, "~", fill_col))
    if (print == TRUE){print(my_formula)}
    # 1. Create fake statistical test data
    fake_stat_test <- metrics_df %>%
            slice(rep(1:n(), each = 2)) %>%
        # Add metrics for lower and upper bounds
        mutate(
            !!sym(y_col) := ifelse(row_number() %% 2 == 1, !!sym(lower_col), !!sym(upper_col))
        )
        # Replicate each row n times
        # slice(rep(1:n(), each = n_replications)) %>%
        # # Add random metrics (these could be parameterized if needed)
        # mutate(
        #     !!sym(y_col) := runif(n(), min = lower_roc_auc, max = upper_roc_auc)
        #     # !!sym(paste0("std_", metric_name)) := runif(n(), min = 0.003, max = 0.02),
        #     # !!sym(paste0("lower_", metric_name)) := runif(n(), min = 0.49, max = 0.55),
        #     # !!sym(paste0("upper_", metric_name)) := runif(n(), min = 0.55, max = 0.61)
        # )

    
    if (print == TRUE){
      message("fake_stat_test")
      print(fake_stat_test)}

    fake_stat_test  <- fake_stat_test %>% 
        # Group and perform t-test
        group_by(!!!syms(columns_to_group)) %>%
        rstatix::t_test(my_formula, comparisons = runs_to_compare) %>%
        adjust_pvalue(method = "bonferroni") %>%
        add_significance() %>%
        add_combination() %>%
        rename(p_value = p, p_value_bonf_diag = p.adj) %>%
        mutate(p_value_bonf_diag = 1,
        p_value_fdr_diag = 1,
        p_value_bonf_all = 1,
        p_value_fdr_all = 1,
        sig = "ns",
        sig_bonf_diag = "ns",
        sig_bonf_all = "ns",
        sig_fdr_diag = "ns",
        sig_fdr_all = "ns") %>%
        select(-c(n1, n2, statistic, df))

    
    if (print == TRUE){
      message("fake_stat_test")
      print(fake_stat_test)}
    
    if (print == TRUE){
      message("emp_pvalues_df")
      print(emp_pvalues_df)}
    # 2. Process empirical p-values
    processed_emp_pvalues <- emp_pvalues_df %>%
        rename_groups() %>%
        add_combination() %>%
        select(!-any_of(colnames(fake_stat_test)))

    if (print == TRUE){
      message("processed_emp_pvalues")
      print(processed_emp_pvalues)}
    
    # print(fake_stat_test)
    # 3. Add x-y positions for plotting
    fake_stat_test <- fake_stat_test %>% 
        filter(combination %in% processed_emp_pvalues$combination) %>% 
        add_xy_position(x = x_col, dodge = 0.8, step.increase = step_increase, fun = fun_to_apply)
        # add_xy_position(fun = fun_to_apply, dodge = 0.8, step.increase = step_increase)

    if (print == TRUE){
      message("fake_stat_test")
      print(fake_stat_test)}

    # 4. Update fake stats with empirical p-values
    new_stat_test <- fake_stat_test %>%
        rows_update(processed_emp_pvalues, by = "combination", unmatched = "ignore")

    if (print == TRUE){
      message("new_stat_test")
      print(new_stat_test)}
    return(new_stat_test)
}


excel_filename  <- "eirpsy_tables.xlsx"
# #-
# eirpsy_theme <- theme_bw()+
# theme(
#   text = element_text(size=18, color = "black"),
#   plot.title = element_text(size=18, hjust = 0.5, color = "black"),
#   axis.text = element_text(color = "black"),
#   axis.title = element_text(color = "black"),
#   legend.text = element_text(color = "black"),
#   legend.title = element_text(color = "black")
# )
# # set theme
# theme_set(eirpsy_theme)

eirpsy_theme  <- theme_bw() + 
theme(
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  plot.title = element_text(hjust = 0.5),
  text=element_text(size=16),
  panel.border = element_rect(linewidth = 1.5),
  # center x label under the tick
    # axis.text.x = element_text(hjust = 0.5),
  axis.ticks.x = element_blank(),
  # change all text color to black
  axis.text = element_text(color = "black")
  # axis.text.x = element_text(color = "black"),
)

# theme_set(eirpsy_theme)

tocount  <- function(dataframe){
  dataframe  %>% 
  count(Model, GLN, Interaction, Label, Features, Features_set, CV) %>% 
  print(n = Inf)
}
diagnoses <- c("ADHD", "ASD", "MDD", "SCZ", "BIP")

fold_eir_vs_logreg  <- "04_EIR_vs_logreg_with_PGS_and_FGRS"

cols_to_group  <- c("ID", "Model", "GLN", "Interaction", "Features", "Features_set", "Diagnosis", "True_Label", "CV", "Split_trained", "Split_eval")

cols_to_select  <- c("ID", "Model","GLN","Interaction","Features","Features_set","Diagnosis","True_Label", "CV", "Split_trained", "Split_eval", "Score_Case_mean", "Score_Control_mean", "Score_Case_std", "Score_Control_std", "prevalence", "m_case")

all_models  <- c(
  "EIR: PGS",
  "EIR: PGS + inter",
  "EIR: PGS_FGRS",
  "EIR: PGS_FGRS + inter",
  "EIR: FGRS",
  "EIR: FGRS + inter",
  "EIR: GLN",
  "EIR: GLN_FGRS",
  "EIR: GLN_FGRS + inter",
  "EIR: GLN_PGS",
  "EIR: GLN_PGS + inter",
  "EIR: GLN_PGS_FGRS",
  "EIR: GLN_PGS_FGRS + inter",
  "EIR: GLN_bigstatsr",
  "EIR: GLN_bigstatsr + inter",
  "EIR: GLN_bigstatsr_PGS",
  "EIR: GLN_bigstatsr_PGS + inter",
  # "EIR: GLN_bigstatsr_FGRS",
  "EIR: GLN_bigstatsr_FGRS + inter",
  "EIR: GLN_bigstatsr_PGS_FGRS",
  "EIR: GLN_bigstatsr_PGS_FGRS + inter",
  "EIR: bigstatsr",
  "EIR: bigstatsr + inter",
  "EIR: bigstatsr_PGS",
  "EIR: bigstatsr_FGRS",
  "EIR: bigstatsr_PGS + inter",
  "EIR: bigstatsr_PGS_FGRS",
  "EIR: bigstatsr_PGS_FGRS + inter",
  "Logreg: FGRS",
  "Logreg: GLN",
  "Logreg: GLN_PGS",
  "Logreg: GLN_PGS + inter",
  "Logreg: GLN_PGS_FGRS",
  "Logreg: GLN_PGS_FGRS + inter",
  "Logreg: GLN_bigstatsr_PGS_FGRS",
  "Logreg: GLN_bigstatsr_PGS_FGRS + inter",
  "Logreg: GLN_bigstatsr",
  "Logreg: GLN_bigstatsr + inter",
  "Logreg: PGS",
  "Logreg: PGS_FGRS",
  "Logreg: PGS_FGRS + inter",
  "Logreg: bigstatsr",
  "Logreg: bigstatsr_PGS",
  "Logreg: bigstatsr_PGS + inter",
  "Logreg: bigstatsr_FGRS",
  "Logreg: bigstatsr_FGRS + inter",
  "Logreg: bigstatsr_PGS_FGRS",
  "Logreg: bigstatsr_PGS_FGRS + inter",
  "bigstatsr: bigstatsr"
)

# GLN_bigstatsr_FGRS missing as a run
# fix bigstatsr position in Logreg


all_labels  <- unique(str_extract(all_models, "(?<=: ).*")); all_labels
all_labels  <- all_labels[!(!str_detect(all_labels, "_") & str_detect(all_labels, "inter"))]; all_labels

all_labels_ordered <- c(
  "GLN", "bigstatsr", "PGS", "FGRS",
  "PGS_FGRS",
  "GLN_FGRS", "GLN_FGRS + inter", 
  "GLN_PGS", "GLN_PGS + inter", 
  "GLN_bigstatsr", "GLN_bigstatsr + inter", 
  "bigstatsr_PGS", "bigstatsr_PGS + inter",
  "bigstatsr_FGRS", "bigstatsr_FGRS + inter",
  "PGS_FGRS + inter",
  "GLN_PGS_FGRS", "GLN_PGS_FGRS + inter",
  "GLN_bigstatsr_PGS", "GLN_bigstatsr_PGS + inter", 
  "GLN_bigstatsr_FGRS + inter", 
  "GLN_bigstatsr_PGS_FGRS", "GLN_bigstatsr_PGS_FGRS + inter", 
  "bigstatsr_PGS_FGRS", "bigstatsr_PGS_FGRS + inter"
)

all_labels_ordered <- list(
  one = c("GLN", "bigstatsr", "PGS", "FGRS"),
  two = c("PGS_FGRS", "PGS_FGRS + inter",
  "GLN_FGRS", "GLN_FGRS + inter", 
  "GLN_PGS", "GLN_PGS + inter", 
  "GLN_bigstatsr", "GLN_bigstatsr + inter", 
  "bigstatsr_PGS", "bigstatsr_PGS + inter",
  "bigstatsr_FGRS", "bigstatsr_FGRS + inter"),

  three_one_geno = c("GLN_PGS_FGRS", "GLN_PGS_FGRS + inter",
    "bigstatsr_PGS_FGRS", "bigstatsr_PGS_FGRS + inter"),
  three_two_geno = c("GLN_bigstatsr_PGS", "GLN_bigstatsr_PGS + inter", 
  "GLN_bigstatsr_FGRS", "GLN_bigstatsr_FGRS + inter"),

  four = c("GLN_bigstatsr_PGS_FGRS", "GLN_bigstatsr_PGS_FGRS + inter")

)

andrew_comp  <- c(all_labels_ordered[["one"]],   "GLN_bigstatsr", "GLN_bigstatsr + inter", all_labels_ordered[["three_one_geno"]], all_labels_ordered[["three_two_geno"]], all_labels_ordered[["four"]])
# create dictionary style

andrew_comp_loc <- andrew_comp[!andrew_comp %in% c("GLN_bigstatsr_PGS", "GLN_bigstatsr_FGRS", "GLN_bigstatsr_PGS + inter", "GLN_bigstatsr_FGRS + inter")]





```

## Hypertune AUCs
```{r, eval = TRUE}
# ----------------------------------
# Hypertuning AUCs
# ----------------------------------
if (working_env == "genomedk"){
  hypertuning_aucs  <- hypertune_preds_hypers %>% 
  select(ID:Diagnosis, Hypercomb, Valid, all_of(hypers_changed)) %>% 
  group_by(Diagnosis, Hypercomb, across(all_of(hypers_changed))) %>% 
  roc_auc(True_Label, Score_Case, event_level = "second") %>% 
  rename_with(~str_replace(., "_input_gln_config", ""), contains("_input_gln_config")) %>% 
  rename(roc_auc = .estimate)
  write_csv(hypertuning_aucs, file.path(current_dir, "data", "01_hyperparameter_tuning", "hypertuning_aucs.csv"))

} else if(working_env == "local"){
  hypertuning_aucs  <- read_csv(file.path("data", "hypertuning_aucs.csv"))
}


library(openxlsx)




library(knitr)
library(kableExtra)

table_best_hypers  <- hypertuning_aucs %>% 
group_by(Diagnosis) %>% 
filter(roc_auc == max(roc_auc)) %>% 
rename(AUROC = roc_auc) %>% 
select(-.metric, -.estimator) %>%
rename(Hypercomb_number = Hypercomb) %>% 
ungroup()

hypertuning_aucs %>% count(Hypercomb) %>% print(n=Inf)

```

## Hypertune Parallel coordinates plot

```{r, eval = TRUE}
# ----------------------------------
# Parallel coordinate plot
# ----------------------------------

create_parallel_plot <- function(df, diagnosis, metric_name) {
  plot <- df %>%
    filter(.metric == metric_name) %>%
    filter(Diagnosis == diagnosis) %>%
    plot_ly(
      type = 'parcoords',
      line = list(color = ~roc_auc, colorscale = 'Viridis', showscale = TRUE),
      dimensions = list(
        list(range = range(df$kernel_width),
             label = 'kernel_width', values = ~kernel_width),
        list(range = range(df$channel_exp_base),
             label = 'channel_exp_base', values = ~channel_exp_base),
        list(range = range(df$rb_do),
             label = 'rb_do', values = ~rb_do),
        list(range = range(df$na_augment_perc),
             label = 'na_augment_perc', values = ~na_augment_perc),
        list(range = range(df$roc_auc),
             label = metric_name, values = ~roc_auc)
      )
    ) %>%
    layout(
      title = glue("Hyperparameter tuning for {diagnosis}"),
      font = list(family = "Arial", size = 12)
    )
  
  # Display the plot
  plot
}

#-
create_parallel_plot <- function(df, diagnosis, column_name = "avg", color_scale = "Jet") {
  
  # Convert all relevant columns to factors and then to numeric sequences for plotting
  df$kernel_width_cat <- as.numeric(as.factor(df$kernel_width))
  df$channel_exp_base_cat <- as.numeric(as.factor(df$channel_exp_base))
  df$rb_do <- as.numeric(as.factor(df$rb_do))
  df$na_augment_perc <- as.numeric(as.factor(df$na_augment_perc))
  
  # Create mappings of the factor levels for labeling
  kernel_width_levels <- levels(as.factor(df$kernel_width))
  channel_exp_base_levels <- levels(as.factor(df$channel_exp_base))
  rb_do_levels <- levels(as.factor(df$rb_do))
  na_augment_perc_levels <- levels(as.factor(df$na_augment_perc))
  
  plot <- df %>%
    plot_ly(
      type = 'parcoords',
      line = list(color = ~get(column_name), colorscale = color_scale, showscale = TRUE),
      dimensions = list(
        list(range = c(1, length(kernel_width_levels)),
             tickvals = 1:length(kernel_width_levels),
             ticktext = kernel_width_levels,
             label = 'Kernel width', values = ~kernel_width_cat),
        list(range = c(1, length(channel_exp_base_levels)),
             tickvals = 1:length(channel_exp_base_levels),
             ticktext = channel_exp_base_levels,
             label = 'Channel exp base', values = ~channel_exp_base_cat),
        list(range = c(1, length(rb_do_levels)),
             tickvals = 1:length(rb_do_levels),
             ticktext = rb_do_levels,
             label = 'Residual Block Dropout', values = ~rb_do),
        list(range = c(1, length(na_augment_perc_levels)),
             tickvals = 1:length(na_augment_perc_levels),
             ticktext = na_augment_perc_levels,
             label = 'NA Augment Perc', values = ~na_augment_perc),
        list(range = range(df$roc_auc),
             label = 'AUROC', values = ~roc_auc)
      )
    ) %>%
    plotly::layout(
      title = glue("Hyperparameter tuning for {diagnosis}"),
      font = list(family = "Arial", size = 18)
    )

  # Display the plot
  plot
}

library(htmlwidgets)

library(plotly)
parallel_adhd  <- create_parallel_plot(hypertuning_aucs, "ADHD", "roc_auc")
saveWidget(parallel_adhd, file.path(figs_path,"parallel_plot_adhd.html"))
parallel_asd  <- create_parallel_plot(hypertuning_aucs, "ASD", "roc_auc")
saveWidget(parallel_asd, file.path(figs_path,"parallel_plot_asd.html"))
parallel_bip  <- create_parallel_plot(hypertuning_aucs, "BIP", "roc_auc")
saveWidget(parallel_bip, file.path(figs_path,"parallel_plot_bip.html"))
parallel_mdd  <- create_parallel_plot(hypertuning_aucs, "MDD", "roc_auc")
saveWidget(parallel_mdd, file.path(figs_path,"parallel_plot_mdd.html"))
parallel_scz  <- create_parallel_plot(hypertuning_aucs, "SCZ", "roc_auc")
saveWidget(parallel_scz, file.path(figs_path,"parallel_plot_scz.html"))






# add_sheet(excel_filename, sheetName = "Table S2", line = "Table S2: Best hyperparameter combinations of the EIR GLN models for ADHD, ASD, BIP, MDD, SCZ", 
# table  = table_best_hypers)


```

## Hypertune pairwise relationship of hyperparameters vs ROC-AUC
```{r, eval = TRUE}

rename_df  <- tribble(
  ~old, ~new,
  "channel_exp_base", "Channel Exp Base",
  "kernel_width", "Kernel Width",
  "rb_do", "Residual Block Dropout",
  "na_augment_perc", "NA Augment Perc",
  "roc_auc", "AUROC"
)

hypertuning_aucs  <- hypertuning_aucs %>% 
rename_with(~rename_df$new[match(., rename_df$old)], all_of(rename_df$old))

# Create a named vector for mapping old names to new names
hypers_changed_long <- c()
for (hyper in hypers_changed_short) {
  idx <- match(hyper, rename_df$old)
  if (!is.na(idx)) {
    hypers_changed_long <- c(hypers_changed_long, rename_df$new[idx])
  } else {
    hypers_changed_long <- c(hypers_changed_long, hyper)
  }
}


p_ggduo_recon  <- ggduo(data = hypertuning_aucs, columnsX = hypers_changed_long, columnsY = c("AUROC"), types = list(continuous = "smooth_lm")
      ) +
  theme_bw()+
  ggtitle(glue("Pairwise relationships between Hyperparameters and metrics"))+
  theme(plot.title = element_text(hjust = 0.5),
  # strip.background = element_rect(fill = "transparent"),
  # strip.background = element_blank(),
  axis.text.x = element_text(angle = 45, hjust = 1),
  text = element_text(size = 20))+
  
  ; p_ggduo_recon


hyper  <- "channel_exp_base"
diagnosis  <- "SCZ"

p_scatter <- ggplot(data = hypertuning_aucs %>% filter(Diagnosis == diagnosis), aes(x = !!sym(hyper), y = roc_auc)) +
  geom_point(alpha = 0.7) + # Scatter points
  geom_smooth(method = "lm", se = TRUE, color = "blue") + # Regression line
  stat_poly_eq(
    aes(label = paste(after_stat(eq.label), after_stat(p.value.label), sep = "~~~")),
    formula = y ~ x, parse = TRUE,
    color = "red"
  ) +
  theme_bw() +
  ggtitle(glue("Relationship in {diagnosis} between {hyper} and ROC-AUC")) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 20)
  );p_scatter



#-
create_equations <- function(data, mapping, ...) {
  # Get the x variable name using rlang
  x_var <- as_name(mapping$x)
  x_unique <- sort(unique(data[[x_var]]))
  
  ggplot(data = data, mapping = mapping) +
    geom_point(alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "blue") +
    stat_poly_eq(
      aes(label = paste(..eq.label.., ..p.value.label.., sep = "~~~")),
      formula = y ~ x, parse = TRUE, color = "red"
    ) +
    scale_x_continuous(breaks = x_unique) +
    facet_grid(Diagnosis ~ .)
}


p_pairwise_relationship <- ggduo(
  data = hypertuning_aucs,
  columnsX = hypers_changed_long,
  columnsY = c("AUROC"),
  types = list(continuous = create_equations)
) +
  ggtitle(glue("Pairwise relationships between each hyperparameter and AUROC")) +
  theme_bw()+
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 20)
  );p_pairwise_relationship

ggsave(file.path(figs_path, "FigS2_hyper_pairwise_relationships.pdf"), p_pairwise_relationship, width = 35, height = 35, units = "cm")




hypertuning_aucs %>% 
group_by(Diagnosis) %>%
summarize(min_auc = min(roc_auc), max_auc = max(roc_auc))




```


# GLN 5-CV performances
```{r, eval = TRUE}

if(working_env == "genomedk"){
  pres_gln_valids  <- readRDS(file.path(current_dir, "data", "02_GLN_vs_bigstatsr", "eir_GLNonly_preds_valid_5CV.rds"))


  valids_aucs  <- pres_gln_valids %>% 
  mutate(
  Valid = as.factor(Valid),
  True_Label = as.factor(True_Label)) %>% 
  group_by(Diagnosis, Valid) %>% 
  roc_auc(True_Label, Score_Case, event_level = "second")

  write_csv(valids_aucs, file.path(current_dir, "data", "02_GLN_vs_bigstatsr", "valids_aucs.csv"))
} else if(working_env == "local"){
  message(glue("Working locally, reading file from {file.path('data', 'valids_aucs.csv')}"))
  valids_aucs  <- read_csv(file.path("data", "valids_aucs.csv"))
}


p_valid_aucs  <- valids_aucs %>% 
mutate(Valid = str_replace(Valid, "valid_", "")) %>%
ggplot(., aes(x = Diagnosis, y = .estimate, fill = Diagnosis))+
geom_boxplot()+
# add jitter points with black circumference
geom_point(aes(fill = Diagnosis), shape = 21, size = 3, color = "black", alpha = 1)+
labs(y = "AUROC", x = "Diagnosis", title = "")+
# scale_fill_manual(values = cbp1[c(2,3,1)], name = "SNP Set")+
# scale_color_manual(values = cbp1[c(2,3,1)], name = "")+
# theme_bw()+
# remove legend of point
guides(fill = guide_legend(override.aes = list(shape = NA)))+
# geom_text(aes(label = Valid), size = 3, vjust = -0.2, hjust = -0.5)+
stat_summary(fun = mean, geom = "text", aes(label = paste("Avg:", round(after_stat(y), 3))), vjust = -1, color = "black", size = 7)+
# add horizontal dashed line at y = 0.5
geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 1.5)+
my_theme; p_valid_aucs

ggsave(file.path(figs_path, "FigS3_gln_valid_aucs.pdf"), p_valid_aucs, width = 35, height = 30, units = "cm")

# add_sheet(excel_filename, sheetName = "Table S3", line = "Table S3: Classification performance of the EIR GLN models with best hyperparameters on 5 validation folds ADHD, ASD, BIP, MDD, SCZ", table =valids_aucs %>% 
# mutate(Valid = str_replace(Valid, "valid_", "")) %>% 
# rename(AUROC = .estimate) %>%
# select(Diagnosis, Valid, AUROC))






```

# Format metrics_repl
This dataframe contains the 10,000 bootstrapped AUROCs for each Diagnosis-Model-Feature-Split_eval combination apart from the ADHD one, since I made a mistake when I asked to export the data
```{r}

metrics_repl  <- readRDS(file.path(current_dir, "data", "metrics_full_gln_big_eir_logreg_sameboots.rds"))

# Ugly but necessary reformatting


metrics_repl  <- metrics_repl %>% 
mutate(Split_trained = "train_valid") %>% 
mutate(Features = case_when(
  Model == "EIR" & Features == "genotype" ~ "GLN_on_geno",
  TRUE ~ Features
)) %>% 
mutate(Features = case_when(
  Model == "bigstatsr" & Features == "genotype" ~ "bigstatsr_on_geno",
  TRUE ~ Features
)) %>%
mutate(Model_type = case_when(
  Model == "EIR" ~ "DL",
  Model %in% c("Logreg", "bigstatsr") ~ "Linear",
  TRUE ~ NA_character_
))  %>% 
mutate(Split_eval = case_when(
  Split_eval == "iPSYCH2015i" ~ "iPSYCH2",
  Split_eval == "test" ~ "test_iPSYCH1",
  TRUE ~ Split_eval
)) %>% 
mutate(Split_eval = fct_relevel(Split_eval, c("valid","test_iPSYCH1", "iPSYCH2"))) %>% 
mutate(Model_genotype = case_when(
  Model == "EIR" & Features == "GLN_on_geno" ~ "GLN",
  Model == "bigstatsr" & Features == "bigstatsr_on_geno" ~ "bigstatsr",
  TRUE ~ NA_character_
)) %>% 
mutate(Model_genotype = fct_relevel(Model_genotype, c("GLN", "bigstatsr"))) %>% 
mutate(Features = case_when(
Features == "GLN+PGS+FGRS+bigstatsr" ~ "GLN+bigstatsr+PGS+FGRS",
TRUE ~ Features
)) %>% 
mutate(Features = case_when(
Features == "GLN on geno" ~ "GLN_on_geno",
Features == "bigstatsr on geno" ~ "bigstatsr_on_geno",
TRUE ~ Features
))


# Select only the R2 computed from coef_to_liab(R2(x,y)) as in Florian's way
metrics_repl  <- metrics_repl %>% 
select(-contains("nag"), -contains("_ABC"), -ends_with("r2_lin")) %>% 
select(-r2_lin_replicates) %>% 
rename_with(~str_remove(., "_lin_liab_Florian"))


# ----------------------------------
# Formatting
# ----------------------------------


metrics_big_repl  <- metrics_repl %>% 
filter(Features == "bigstatsr_on_geno") %>% 
mutate(Features = "bigstatsr")  

metrics_big_eir_repl  <- metrics_repl %>%
filter(Features == "bigstatsr" & Interaction == "Inter")


metrics_gln_repl  <- metrics_repl %>%
filter(Features == "GLN_on_geno") %>% 
mutate(Features = "GLN")

metrics_gln_logreg_repl  <- metrics_repl %>% 
filter(Features == "GLN" & Model == "Logreg" & CV == "5-CV")

metrics_pgs_fgrs_eir_repl  <- metrics_repl %>% 
filter(CV == "5-CV") %>% 
filter(str_detect(Features, "PGS") | str_detect(Features, "FGRS")) %>% 
filter(!str_detect(Features, "\\+")) %>%
filter(Interaction == "Inter")

# metrics_pgs_fgrs_eir %>% count_cols()
metrics_pgs_fgrs_logreg_repl  <- metrics_repl %>% 
filter(CV == "5-CV") %>% 
filter(str_detect(Features, "PGS") | str_detect(Features, "FGRS")) %>% 
filter(!str_detect(Features, "\\+")) %>%
filter(Model_type == "Linear")

# metrics_pgs_fgrs_logreg %>% count_cols()

metrics_multiple_inter_repl  <- metrics_repl %>%
filter(CV == "5-CV") %>% 
filter(str_detect(Features, "\\+") & Interaction == "Inter")



metrics_adhd_asd_mdd_inter_repl  <- metrics_repl %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("ADHD", "ASD", "MDD")) %>% 
filter(Interaction == "Inter") %>% 
# filter(!Features %in% c("GLN", "PGS", "FGRS")) %>% 
filter(str_detect(Features, "\\+"))

# For BIP and SCZ I want to keep only the models run without interaction because probably because of the small sample size, the additional term impacted negatively on the performance
metrics_scz_bip_nointer_repl  <- metrics_repl %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("SCZ", "BIP")) %>% 
filter(Interaction == "No-Inter" ) %>% 
filter(str_detect(Features, "\\+"))


metrics_scz_bip_nointer_repl_linear  <- metrics_repl %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("SCZ", "BIP")) %>% 
filter(Interaction == "No-Inter" & Integration == "Linear") %>% 
filter(str_detect(Features, "\\+"))

metrics_scz_bip_nointer_repl_DL  <- metrics_repl %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("SCZ", "BIP")) %>% 
filter(Interaction == "Inter" & Integration == "DL") %>% 
filter(str_detect(Features, "\\+"))


# metrics_multiple_inter_repl <- bind_rows(metrics_adhd_asd_mdd_inter_repl, metrics_scz_bip_nointer_repl)
metrics_multiple_inter_repl <- bind_rows(metrics_adhd_asd_mdd_inter_repl, metrics_scz_bip_nointer_repl_linear, metrics_scz_bip_nointer_repl_DL)


to_concatenate  <- c("Diagnosis", "Model", "Features", "Split_eval", "Interaction", "CV", "Split_trained")

metrics_dl_vs_linear_repl  <- bind_rows(
  metrics_big_repl,
  # metrics_big_eir_repl,
  metrics_gln_repl,
  # metrics_gln_logreg_repl,
  metrics_pgs_fgrs_eir_repl,
  metrics_pgs_fgrs_logreg_repl,
  metrics_multiple_inter_repl
) %>% 
filter_out_ADHD() %>% 
mutate(Concatenated = paste(!!!syms(to_concatenate), sep = "_")) %>% 
filter(Split_eval != "valid")
# There are no duplicates


```


# Format metrics_avg
```{r}

metrics_avg  <- metrics_repl %>% 
select(-contains("replicates")) %>% 
select(-c(Model_type, Model_genotype)) %>% 
pivot_longer(cols = -c(Diagnosis, Model,GLN, Interaction, Features, CV, Split_trained, Split_eval, Integration, Label), names_to = "metric", values_to = "estimate") %>% 
select(Diagnosis, Model,GLN, Interaction, Features, CV, Split_trained, Split_eval, metric, estimate, Integration, Label)

metrics_avg_orig  <- metrics_avg

metrics_avg  <- metrics_avg %>% 
mutate(Split_trained = "train_valid")

metrics_avg %>% 
filter(Model == "Logreg" & Features == "bigstatsr") %>% count_cols()
# There are 4 but I don't need the last 2 and I will replace the first 2 with the ones from bigstatsr

metrics_avg %>% 
filter(Model == "bigstatsr") %>% count_cols()

metrics_avg <- metrics_avg %>% 
filter(!(Model == "Logreg" & Features == "bigstatsr")) %>% 
# Convert "genotype" in EIR to GLN. This is to plot them
mutate(Features = case_when(
  Model == "EIR" & Features == "genotype" ~ "GLN_on_geno",
  TRUE ~ Features
)) %>% 
mutate(Features = case_when(
  Model == "bigstatsr" & Features == "genotype" ~ "bigstatsr_on_geno",
  TRUE ~ Features
)) %>%
mutate(Model_type = case_when(
  Model == "EIR" ~ "DL",
  Model %in% c("Logreg", "bigstatsr") ~ "Linear",
  TRUE ~ NA_character_
))  %>% 
mutate(Split_eval = case_when(
  Split_eval == "iPSYCH2015i" ~ "iPSYCH2",
  Split_eval == "test" ~ "test_iPSYCH1",
  TRUE ~ Split_eval
)) %>% 
mutate(Split_eval = fct_relevel(Split_eval, c("valid","test_iPSYCH1", "iPSYCH2"))) %>% 
mutate(Model_genotype = case_when(
  Model == "EIR" & Features == "GLN_on_geno" ~ "GLN",
  Model == "bigstatsr" & Features == "bigstatsr_on_geno" ~ "bigstatsr",
  Model == "bigstatsr+PCs" & Features == "bigstatsr+PCs" ~ "bigstatsr+PCs",
  Model == "bigstatsr+sex+age+PCs" & Features == "bigstatsr+sex+age+PCs" ~ "bigstatsr+sex+age+PCs",
  TRUE ~ NA_character_
)) %>% mutate(Model_genotype = fct_relevel(Model_genotype, c("GLN", "bigstatsr")))

metrics_avg  <- metrics_avg %>% 
  mutate(metric = case_when(
    metric == "avg_auc" ~ "avg_roc_auc",
    metric == "sd_auc" ~ "std_roc_auc",
    metric == "sd_r2" ~ "std_r2",
    metric == "lower_auc" ~ "lower_roc_auc",
    metric == "upper_auc" ~ "upper_roc_auc",
    TRUE ~ metric
  ))

```


# P-values
## p-values each model-feature iPSYCH2 - iPSYCH1 
Compute the empirical p-values of the difference in performance between iPSYCH2 and iPSYCH1 for each Diagnosis-Model-Feature combination.

```{r}

# ----------------------------------
# Same feature same model iPSYCH2 - iPSYCH1 difference
# Question: Is the Model-Feature's performance in iPSYCH2 higher than in iPSYCH1?
# ----------------------------------

# I will need these for all the models

to_compare <- "same_feature_same_model_different_eval"

comparison_params <- set_comparison_parameters(to_compare)

res_tabs_same_feature_same_model_different_eval <- metrics_dl_vs_linear_repl %>% 
# select(GLN, Interaction) %>% 
filter(Split_eval %in% c("test_iPSYCH1", "iPSYCH2"))
date
folder_name <- glue("emp_pvalues_{date}_{to_compare}")
filename  <- "auc_pvalues_same_feature_same_model_different_eval.rds"
# remove file if it exists
# file.remove(file.path(current_dir, folder_name, filename))


if(!file.exists(file.path(final_folder, folder_name, filename))){
auc_pvalues_same_feature_same_model_different_eval_tabs <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = res_tabs_same_feature_same_model_different_eval  %>% filter(!Features %in% c("GLN","GLN_on_geno","bigstatsr","bigstatsr_on_geno")),
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Split_eval",
  folder_name = folder_name
); auc_pvalues_same_feature_same_model_different_eval_tabs

auc_pvalues_same_feature_same_model_different_eval_gln_big <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = res_tabs_same_feature_same_model_different_eval %>% filter(Model_genotype %in% c("GLN", "bigstatsr")),
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Split_eval",
  folder_name = folder_name
); auc_pvalues_same_feature_same_model_different_eval_gln_big

  auc_pvalues_same_feature_same_model_different_eval <- bind_rows(auc_pvalues_same_feature_same_model_different_eval_tabs, auc_pvalues_same_feature_same_model_different_eval_gln_big)
  saveRDS(auc_pvalues_same_feature_same_model_different_eval, file.path(folder_name, filename))

} else {
  auc_pvalues_same_feature_same_model_different_eval <- readRDS(file.path(final_folder, folder_name, filename))
}


filename  <- "r2_pvalues_same_feature_same_model_different_eval.rds"

if(!file.exists(file.path(final_folder, folder_name, filename))){
  r2_pvalues_same_feature_same_model_different_eval_tabs <- compute_pvalues_comparisons(
    to_compare = to_compare,
    data = res_tabs_same_feature_same_model_different_eval  %>% filter(!Features %in% c("GLN","GLN_on_geno","bigstatsr","bigstatsr_on_geno")),
    columns_to_group = comparison_params$columns_to_group,
    runs_to_compare = comparison_params$runs_to_compare,
    filter_column = "Split_eval",
    metric = "r2",
    save_plots = TRUE,
    folder_name = folder_name
  )

  r2_pvalues_same_feature_same_model_different_eval_gln_big <- compute_pvalues_comparisons(
    to_compare = to_compare,
    data = res_tabs_same_feature_same_model_different_eval %>% filter(Model_genotype %in% c("GLN", "bigstatsr")),
    columns_to_group = comparison_params$columns_to_group,
    runs_to_compare = comparison_params$runs_to_compare,
    filter_column = "Split_eval",
    metric = "r2",
    save_plots = TRUE,
    folder_name = folder_name
    )

  r2_pvalues_same_feature_same_model_different_eval <- bind_rows(r2_pvalues_same_feature_same_model_different_eval_tabs, r2_pvalues_same_feature_same_model_different_eval_gln_big)

  saveRDS(r2_pvalues_same_feature_same_model_different_eval, file.path(final_folder, folder_name, filename))
} else {
  r2_pvalues_same_feature_same_model_different_eval <- readRDS(file.path(final_folder, folder_name, filename))
}

```


## p-values GLN vs bigstatsr on test iPSYCH1

```{r}

#-
runs_compare_eir_vs_logreg  <- c(
  "GLN_on_geno", 
  "bigstatsr_on_geno", 
  "GLN+bigstatsr",
  "PGS", 
  "FGRS", 
  "PGS+FGRS",
  "GLN+PGS+FGRS",
  "bigstatsr+PGS+FGRS",
  "GLN+bigstatsr+PGS+FGRS"
)

features_compare_gln_vs_big_test_repl  <- c(
  "GLN_on_geno", 
  "bigstatsr_on_geno"
)

labels_gln_big  <- c(
  "EIR: genotype", 
  "bigstatsr: genotype"
)
to_compare <- "GLN_vs_bigstatsr"

comparison_params <- set_comparison_parameters(to_compare)
columns_to_group <- comparison_params$columns_to_group
runs_to_compare <- comparison_params$runs_to_compare
runs_to_compare_list <- comparison_params$runs_to_compare_list


folder_name <- file.path(final_folder, glue("emp_pvalues_{date}_{to_compare}"))
filename  <- "auc_pvalues_gln_vs_bigstatsr_test.rds"
# delete file if it exists
# if (file.exists(file.path(current_dir, folder_name, filename))){file.remove(file.path(current_dir, folder_name, filename))}
# file.remove(file.path(current_dir, folder_name, filename))
if(!file.exists(file.path(folder_name, filename))){
auc_pvalue_gln_vs_bigstatsr_test  <- compute_pvalues_comparisons(
  data = metrics_repl %>%
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(Split_eval == "test_iPSYCH1") %>%
          filter(CV == "5-CV") %>%
          filter(Features %in% runs_to_compare[[1]]) %>%
          mutate(Features = fct_relevel(Features, runs_to_compare[[1]])),
  to_compare = to_compare,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = list(c("GLN_on_geno", "bigstatsr_on_geno")),
  filter_column = "Features",
  folder_name = folder_name
); auc_pvalue_gln_vs_bigstatsr_test
  saveRDS(auc_pvalue_gln_vs_bigstatsr_test, file.path(folder_name, filename))
} else {
  auc_pvalue_gln_vs_bigstatsr_test <- readRDS(file.path(folder_name, filename))
}

filename  <- "auc_pvalues_gln_vs_bigstatsr_iPSYCH2.rds"
# file.remove(file.path(current_dir, folder_name, filename))
if(!file.exists(file.path(folder_name, filename))){
auc_pvalue_gln_vs_bigstatsr_iPSYCH2 <- compute_pvalues_comparisons(
  data = metrics_repl %>%
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(Split_eval == "iPSYCH2") %>%
          filter(CV == "5-CV") %>%
          filter(Features %in% runs_to_compare[[1]]) %>%
          mutate(Features = fct_relevel(Features, runs_to_compare[[1]])), 
  to_compare = to_compare,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = runs_to_compare,
  filter_column = "Features",
  folder_name = folder_name
); auc_pvalue_gln_vs_bigstatsr_iPSYCH2
  saveRDS(auc_pvalue_gln_vs_bigstatsr_iPSYCH2, file.path(folder_name, filename))
} else {
  auc_pvalue_gln_vs_bigstatsr_iPSYCH2 <- readRDS(file.path(folder_name, filename))
}


auc_pvalues_gln_vs_bigstatsr <- bind_rows(auc_pvalue_gln_vs_bigstatsr_test, auc_pvalue_gln_vs_bigstatsr_iPSYCH2) %>% 
mutate(
            Model1 = str_replace(Model1, "_on_geno", ""),
            Model2 = str_replace(Model2, "_on_geno", "")
        )


aucs_df_for_pvalues  <- metrics_avg %>% 
filter(Features %in% features_compare_gln_vs_big_test_repl) %>% 
    filter(Label %in% labels_gln_big) %>%
    mutate(Label = fct_relevel(Label, labels_gln_big)) %>%
    filter(metric %in% metrics_auc) %>%
    pivot_wider(names_from = metric, values_from = estimate) %>%
    filter_out_ADHD() %>%  
    mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))


#-
# To convert the format of the p-values dataframe to one compatible with stat_pvalue_manual from ggpubr

if(!file.exists(file.path(folder_name, "auc_formatted_pvalues_gln_vs_bigstatsr.rds"))){
  auc_formatted_pvalues_gln_vs_bigstatsr <- process_statistical_tests(
    metrics_df = aucs_df_for_pvalues,
    emp_pvalues_df = auc_pvalues_gln_vs_bigstatsr,
    metric_name = "auc",
    x_col = "Diagnosis",
    fill_col = "Model_genotype",
    columns_to_group = c("Diagnosis", "Split_eval"),
    keep_group_model_order = TRUE
);auc_formatted_pvalues_gln_vs_bigstatsr %>% select(Diagnosis, Split_eval, group1, group2) %>% 
arrange(Split_eval, Diagnosis,group1, group2)

auc_formatted_pvalues_gln_vs_bigstatsr <- auc_formatted_pvalues_gln_vs_bigstatsr %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

saveRDS(auc_formatted_pvalues_gln_vs_bigstatsr, file.path(folder_name, "auc_formatted_pvalues_gln_vs_bigstatsr.rds"))
} else {
  auc_formatted_pvalues_gln_vs_bigstatsr <- readRDS(file.path(folder_name, "auc_formatted_pvalues_gln_vs_bigstatsr.rds"))
}

# ----------------------------------
# Do the same for R2
# ----------------------------------

folder_name <- file.path(final_folder, glue("emp_pvalues_{date}_{to_compare}"))
filename  <- "r2_pvalues_gln_vs_bigstatsr_test.rds"
# file.remove(file.path(current_dir, folder_name, filename))
if(!file.exists(file.path(folder_name, filename))){
r2_pvalue_gln_vs_bigstatsr_test  <- compute_pvalues_comparisons(
  data = metrics_repl %>%
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(Split_eval == "test_iPSYCH1") %>%
          filter(CV == "5-CV") %>%
          filter(Features %in% runs_to_compare[[1]]) %>%
          mutate(Features = fct_relevel(Features, runs_to_compare[[1]])),
  to_compare = to_compare,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = list(c("GLN_on_geno", "bigstatsr_on_geno")),
  filter_column = "Features",
  metric = "r2",
  folder_name = folder_name
); r2_pvalue_gln_vs_bigstatsr_test
  saveRDS(r2_pvalue_gln_vs_bigstatsr_test, file.path(folder_name, filename))
} else {
  r2_pvalue_gln_vs_bigstatsr_test <- readRDS(file.path(folder_name, filename))
}

filename  <- "r2_pvalues_gln_vs_bigstatsr_iPSYCH2.rds"
# file.remove(file.path(current_dir, folder_name, filename))
if(!file.exists(file.path(folder_name, filename))){
r2_pvalue_gln_vs_bigstatsr_iPSYCH2 <- compute_pvalues_comparisons(
  data = metrics_repl %>%
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(Split_eval == "iPSYCH2") %>%
          filter(CV == "5-CV") %>%
          filter(Features %in% runs_to_compare[[1]]) %>%
          mutate(Features = fct_relevel(Features, runs_to_compare[[1]])), 
  to_compare = to_compare,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = runs_to_compare,
  filter_column = "Features",
  metric = "r2",
  folder_name = folder_name
); r2_pvalue_gln_vs_bigstatsr_iPSYCH2
  saveRDS(r2_pvalue_gln_vs_bigstatsr_iPSYCH2, file.path(folder_name, filename))
} else {
  r2_pvalue_gln_vs_bigstatsr_iPSYCH2 <- readRDS(file.path(folder_name, filename))
}


r2_pvalues_gln_vs_bigstatsr <- bind_rows(r2_pvalue_gln_vs_bigstatsr_test, r2_pvalue_gln_vs_bigstatsr_iPSYCH2) %>% 
mutate(
            Model1 = str_replace(Model1, "_on_geno", ""),
            Model2 = str_replace(Model2, "_on_geno", "")
        )


r2_df_for_pvalues  <- metrics_avg %>%  
filter(Features %in% features_compare_gln_vs_big_test_repl) %>% 
    filter(Label %in% labels_gln_big) %>%
    mutate(Label = fct_relevel(Label, labels_gln_big)) %>%
    filter(metric %in% metrics_r2) %>%
    pivot_wider(names_from = metric, values_from = estimate) %>% 
    filter_out_ADHD() %>%  
    # filter(Diagnosis %in% c("ASD", "MDD", "SCZ", "BIP")) %>% 
    mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))


#-

if(!file.exists(file.path(folder_name, "r2_formatted_pvalues_gln_vs_bigstatsr.rds"))){
# To convert the format of the p-values dataframe to one compatible with stat_pvalue_manual from ggpubr
  r2_formatted_pvalues_gln_vs_bigstatsr <- process_statistical_tests(
      metrics_df = r2_df_for_pvalues,
      emp_pvalues_df = r2_pvalues_gln_vs_bigstatsr,
      metric_name = "r2",
      x_col = "Diagnosis",
      fill_col = "Model_genotype",
      columns_to_group = c("Diagnosis", "Split_eval"),
      keep_group_model_order = TRUE
  )

  r2_formatted_pvalues_gln_vs_bigstatsr <- r2_formatted_pvalues_gln_vs_bigstatsr %>% 
  mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

  saveRDS(r2_formatted_pvalues_gln_vs_bigstatsr, file.path(folder_name, "r2_formatted_pvalues_gln_vs_bigstatsr.rds"))
} else {
  r2_formatted_pvalues_gln_vs_bigstatsr <- readRDS(file.path(folder_name, "r2_formatted_pvalues_gln_vs_bigstatsr.rds"))
}




```



# Plot GLN vs bigstatsr on test_iPSYCH1 and iPSYCH2 + drop performance 14/02/25 11:23 works locally
```{r}



source("plotting_functions.R")

#-
bar_width <- 0.9  # Reduce bar width to add space between bars
pvalue_size <- 8
p_auc_gln_vs_big_test <- metrics_avg %>% 
filter(Features %in% features_compare_gln_vs_big_test_repl) %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
filter_out_ADHD() %>% 
plot_barplots_ggpubr(dataframe = .,
  # to_plot = labels_gln_big,
  x_column = "Diagnosis",
  fill_column = "Model_genotype",
  fill_label = "Model",
  title = "",
  x_title = "",
  metrics = metrics_roc_auc,
  y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Split_eval",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right",
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  pvalue_text_size = 12,
  fill_colors = models_colors_palette,
  theme = my_theme
  # bar_width = bar_width
)+

geom_hline(yintercept = 0.5, linetype = "dashed", color = "black")+
coord_cartesian(ylim = c(0.495, 0.66))+
scale_y_continuous(breaks = c(seq(0.5, 1, by = 0.02)))+

stat_pvalue_manual(
  auc_formatted_pvalues_gln_vs_bigstatsr %>% 
  mutate(y.position = case_when(
  Diagnosis == "ADHD" & Split_eval == "test_iPSYCH1" ~ 0.645,
  Diagnosis == "ASD" & Split_eval == "test_iPSYCH1" ~ 0.585,
  Diagnosis == "BIP" & Split_eval == "test_iPSYCH1" ~ 0.60,
  Diagnosis == "MDD" & Split_eval == "test_iPSYCH1" ~ 0.59,
  Diagnosis == "SCZ" & Split_eval == "test_iPSYCH1" ~ 0.615,
  Diagnosis == "ADHD" & Split_eval == "iPSYCH2" ~ 0.615,
  Diagnosis == "ASD" & Split_eval == "iPSYCH2" ~ 0.575,
  Diagnosis == "BIP" & Split_eval == "iPSYCH2" ~ 0.55,
  Diagnosis == "MDD" & Split_eval == "iPSYCH2" ~ 0.56,
  Diagnosis == "SCZ" & Split_eval == "iPSYCH2" ~ 0.565,
  TRUE ~ y.position
  )), label = "p.adj", tip.length = 0.01,
  hide.ns = FALSE,
  size = pvalue_size
  )+
  theme(plot.subtitle = element_text(size = 14));p_auc_gln_vs_big_test
    # labs(subtitle = "test iPSYCH1 H1: AUROC bigstatsr > AUROC GLN,
  # iPSYCH2 H1: AUROC GLN > AUROC bigstatsr")+



p_r2_gln_vs_big_test <- metrics_avg %>% 
filter(Features %in% features_compare_gln_vs_big_test_repl) %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
filter_out_ADHD() %>% 
plot_barplots_ggpubr(dataframe = .,
  # to_plot = labels_gln_big,
  x_column = "Diagnosis",
  fill_column = "Model_genotype",
  fill_label = "Model",
  title = "",
  x_title = "",
  metrics = metrics_r2,
  y_title = "R2",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Split_eval",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right",
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  pvalue_text_size = 12,
  fill_colors = models_colors_palette,
  theme = my_theme
  # bar_width = bar_width
)+
scale_y_continuous(limits = c(0, 0.06))+
labs(fill = "Model")+
stat_pvalue_manual(
  r2_formatted_pvalues_gln_vs_bigstatsr, label = "p.adj", tip.length = 0.01,
  hide.ns = FALSE,
  size = pvalue_size
  )+
  # labs(subtitle = "test iPSYCH1 H1: AUROC bigstatsr > AUROC GLN,
  # iPSYCH2 H1: AUROC GLN > AUROC bigstatsr")+
    theme(
    plot.subtitle = element_text(size = 14)
  );p_r2_gln_vs_big_test


# ggsave(glue("plots_paper_final/figs_{date}/auc_gln_vs_bigstatsr_testiPSYCH1.pdf"), p_auc_gln_vs_big_test)
# ggsave(glue("plots_paper_final/figs_{date}/r2_gln_vs_bigstatsr_testiPSYCH1.pdf"), p_r2_gln_vs_big_test)




```



# Performance drop GLN vs bigstatsr iPSYCH2

## P-values performance drop GLN vs bigstatsr iPSYCH2
Here I compute 2 p-values:
** Within-model p-value: is the performance of each model (GLN and bigstatsr) on test_iPSYCH1 higher than on iPSYCH2?, aka is the performance of each model (GLN and bigstatsr) on iPSYCH2 lower than on test_iPSYCH1?
** Across-model p-value: is the drop in performance of GLN on iPSYCH2 smaller than the drop in performance of bigstatsr on iPSYCH2?

```{r}

# AUROC
trial_metrics_avg <- metrics_avg %>% 
  mutate(metric = case_when(
    metric == "avg_auc" ~ "avg_roc_auc",
    metric == "sd_auc" ~ "std_roc_auc",
    metric == "sd_r2" ~ "std_r2",
    metric == "lower_auc" ~ "lower_roc_auc",
    metric == "upper_auc" ~ "upper_roc_auc",
    TRUE ~ metric
  ))

# Calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the average values
auc_drop_gln_big_repl <- calculate_performance_diff(
  metrics_df = trial_metrics_avg %>% filter_out_ADHD(),
  features_to_compare = features_compare_gln_vs_big_test_repl,
  cols_to_remove = c("CV", "Label", "Model_type", "Split_trained"),
  category_col = "Split_eval",
  metrics_cols = metrics_roc_auc,
  split1 = "test_iPSYCH1",
  split2 = "iPSYCH2"
) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))

# Format the dataframe to calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the bootstrap replicates
column_replicates <- "auc_replicates"
auc_formatted_drop_gln_big_repl  <- metrics_repl %>%          
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(CV == "5-CV") %>%
          filter(Features %in% runs_to_compare[[1]]) %>%
          mutate(Features = fct_relevel(Features, runs_to_compare[[1]])) %>% 
          mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
          select(Diagnosis, Features, Split_eval, Split_trained, CV, all_of(column_replicates)) %>% 
          pivot_wider(names_from = Split_eval, values_from = all_of(column_replicates)) %>% 
          mutate(!!sym(column_replicates) := map2(iPSYCH2, test_iPSYCH1, ~ .x - .y))

# ----------------------------------
# Within-model pvalues
# ----------------------------------


# Get the real p-values for the GLN and bigstatsr models
auc_pvalues_gln_big_iPSYCH2_vs_iPSYCH1 <- auc_pvalues_same_feature_same_model_different_eval %>% 

filter(Features %in% c("GLN","bigstatsr")) %>% 
filter_out_ADHD()



# Format the dataframe to be compatible with geom_text
auc_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1  <- auc_drop_gln_big_repl %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
left_join(auc_pvalues_gln_big_iPSYCH2_vs_iPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))





# ----------------------------------
# R2
# ----------------------------------
r2_drop_gln_big_repl <- calculate_performance_diff(
  metrics_df = trial_metrics_avg %>% filter_out_ADHD(),
  features_to_compare = features_compare_gln_vs_big_test_repl,
  cols_to_remove = c("CV", "Label", "Model_type", "Split_trained"),
  category_col = "Split_eval",
  metrics_cols = metrics_r2,
  split1 = "test_iPSYCH1",
  split2 = "iPSYCH2"
) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))

# Format the dataframe to calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the bootstrap replicates
column_replicates <- "r2_replicates"
r2_formatted_drop_gln_big_repl  <- metrics_repl %>%          
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(CV == "5-CV") %>%
          filter(Features %in% runs_to_compare[[1]]) %>%
          mutate(Features = fct_relevel(Features, runs_to_compare[[1]])) %>% 
          mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
          select(Diagnosis, Features, Split_eval, Split_trained, CV, all_of(column_replicates)) %>% 
          pivot_wider(names_from = Split_eval, values_from = all_of(column_replicates)) %>% 
          mutate(!!sym(column_replicates) := map2(iPSYCH2, test_iPSYCH1, ~ .x - .y))

# ----------------------------------
# Within-model pvalues
# ----------------------------------


# Get the real p-values for the GLN and bigstatsr models
r2_pvalues_gln_big_iPSYCH2_vs_iPSYCH1 <- r2_pvalues_same_feature_same_model_different_eval %>% 
filter(Features %in% c("GLN","bigstatsr")) %>% 
filter_out_ADHD()


# Format the dataframe to be compatible with geom_text
r2_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1  <- r2_drop_gln_big_repl %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
left_join(r2_pvalues_gln_big_iPSYCH2_vs_iPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))


saveRDS(auc_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1, file.path(folder_name, "auc_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1.rds"))
saveRDS(r2_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1, file.path(folder_name, "r2_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1.rds"))



```
## Plot performance drop GLN vs bigstatsr iPSYCH2
```{r}

auc_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1 <- readRDS(file.path(folder_name, "auc_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1.rds"))
r2_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1 <- readRDS(file.path(folder_name, "r2_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1.rds"))


# Barplot
x_column <- "Diagnosis"
y_column <- "diff_avg"
fill_column <- "Model_genotype"
metric_name <- "diff"
bar_width  <- 0.65
y_min  <- -0.05
y_max  <- 0.12

p_auc_perf_drop_gln_vs_big_iPSYCH2  <- ggbarplot(auc_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1, 
  x = x_column, 
  y = y_column, 
  fill = fill_column, 
  position = position_dodge(width = bar_width, preserve = "single"), 
  width = bar_width, 
  add = "", 
  color = "black") +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    # subtitle = "Within-model p-value H1: performance on iPSYCH2 < performance on test_iPSYCH1",
    y = bquote(Delta * AUROC ~ .(suffix_delta_eval)),
    x = "",
    fill = "Model"
  ) +
geom_errorbar(
    aes(
      ymin = !!sym(paste0("lower_", metric_name)),
      ymax = !!sym(paste0("upper_", metric_name)),
      group = interaction(!!sym(x_column), !!sym(fill_column))
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  # scale_y_continuous(breaks = seq(y_min, y_max, by = 0.01), limits = c(y_max, y_min)) +
  # scale_y_reverse(breaks = seq(y_min, y_max, by = 0.01), limits = c(y_max, y_min)) +
  scale_y_reverse(breaks = seq(y_min, y_max, by = 0.02)) +
  my_theme +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = models_colors_palette) +

    geom_text(
    aes(
      x = Diagnosis, 
      y = upper_diff+0.004, 
      # label = format(p_value, nsmall = 2, digits = 2, scientific = FALSE),
      label = p.adj,
      group = Model_genotype
    ),
    position = position_dodge(width = bar_width),
    # vjust = -0.5, # Adjust vertical spacing
    size = pvalue_size
  ) +
  # decrease text size of the subtitle
  theme(
    plot.subtitle = element_text(size = 14)
  );p_auc_perf_drop_gln_vs_big_iPSYCH2



x_column <- "Diagnosis"
p_r2_perf_drop_gln_vs_big_iPSYCH2  <- ggbarplot(r2_perf_drop_gln_big_iPSYCH2_vs_iPSYCH1, 
  x = x_column, 
  y = y_column, 
  fill = fill_column, 
  position = position_dodge(width = bar_width, preserve = "single"), 
  width = bar_width, 
  add = "", 
  color = "black") +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    # subtitle = "Within-model p-value H1: performance on iPSYCH2 < performance on test_iPSYCH1",
    y = bquote(Delta * R^2 ~ .(suffix_delta_eval)),
    x = "",
    fill = "Model"
  ) +
geom_errorbar(
    aes(
      ymin = !!sym(paste0("lower_", metric_name)),
      ymax = !!sym(paste0("upper_", metric_name)),
      group = interaction(!!sym(x_column), !!sym(fill_column))
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  # scale_y_continuous(breaks = seq(y_min, y_max, by = 0.01), limits = c(y_max, y_min)) +
  # scale_y_reverse(breaks = seq(y_min, y_max, by = 0.01), limits = c(y_max, y_min)) +
  scale_y_reverse(breaks = seq(y_min, y_max, by = 0.01)) +
  my_theme +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = models_colors_palette) +
  theme(
    plot.subtitle = element_text(size = 14)
  ) + 

    geom_text(
    aes(
      x = Diagnosis, 
      y = upper_diff+0.003, 
      # label = format(p_value, nsmall = 2, digits = 2, scientific = FALSE),
      label = p.adj,
      group = Model_genotype
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5, # Adjust vertical spacing
    size = pvalue_size
  );p_r2_perf_drop_gln_vs_big_iPSYCH2

ggsave(glue("plots_paper_final/figs_{date}/p_auc_perf_drop_gln_vs_big_iPSYCH2.pdf"), p_auc_perf_drop_gln_vs_big_iPSYCH2)
ggsave(glue("plots_paper_final/figs_{date}/p_r2_perf_drop_gln_vs_big_iPSYCH2.pdf"), p_r2_perf_drop_gln_vs_big_iPSYCH2)


p_gln_big_auc_test  <- p_auc_gln_vs_big_test + theme(legend.position = "none") + p_auc_perf_drop_gln_vs_big_iPSYCH2 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(face = "bold"))


p_gln_big_r2_test  <- p_r2_gln_vs_big_test + theme(legend.position = "none") + p_r2_perf_drop_gln_vs_big_iPSYCH2 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(face = "bold"))



ggsave(file.path(figs_path,"Fig2_gln_big_auc_test_pvalues.pdf"), p_gln_big_auc_test, width = 70, height = 25, units = "cm")
ggsave(file.path(figs_path,"FigS4_gln_big_r2_test_pvalues.pdf"), p_gln_big_r2_test, width = 70, height = 25, units = "cm")



```


# Tabs DL vs Linear on test + iPSYCH2
## Format df
```{r}

# ----------------------------------
# Preparing dataframe NOW 20/01/25 17:57 
# ----------------------------------

#-
metrics_avg <- metrics_avg %>% 
  mutate(metric = case_when(
    metric == "avg_auc" ~ "avg_roc_auc",
    metric == "sd_auc" ~ "std_roc_auc",
    metric == "sd_r2" ~ "std_r2",
    metric == "lower_auc" ~ "lower_roc_auc",
    metric == "upper_auc" ~ "upper_roc_auc",
    TRUE ~ metric
  ))

metrics_big  <- metrics_avg %>% 
filter(Features == "bigstatsr_on_geno") %>% 
mutate(Features = "bigstatsr")

metrics_big_eir  <- metrics_avg %>%
filter(Features == "bigstatsr" & Interaction == "Inter")


metrics_gln  <- metrics_avg %>%
filter(Features == "GLN_on_geno") %>% 
mutate(Features = "GLN")

metrics_gln_logreg  <- metrics_avg %>% 
filter(Features == "GLN" & Model == "Logreg" & CV == "5-CV")

metrics_pgs_fgrs_eir  <- metrics_avg %>% 
filter(CV == "5-CV") %>% 
filter(str_detect(Features, "PGS") | str_detect(Features, "FGRS")) %>% 
filter(!str_detect(Features, "\\+")) %>%
filter(Interaction == "Inter")

# metrics_pgs_fgrs_eir %>% count_cols()

metrics_pgs_fgrs_logreg  <- metrics_avg %>% 
filter(CV == "5-CV") %>% 
filter(str_detect(Features, "PGS") | str_detect(Features, "FGRS")) %>% 
filter(!str_detect(Features, "\\+")) %>%
filter(Model_type == "Linear")

# metrics_pgs_fgrs_logreg %>% count_cols()

metrics_multiple_inter  <- metrics_avg %>%
filter(CV == "5-CV") %>% 
filter(str_detect(Features, "\\+") & Interaction == "Inter")


metrics_adhd_asd_mdd_inter  <- metrics_avg %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("ADHD", "ASD", "MDD")) %>% 
filter(Interaction == "Inter") %>% 
filter(str_detect(Features, "\\+"))

# For BIP and SCZ I want to keep only the models run without interaction because probably because of the small sample size, the additional term impacted negatively on the performance
metrics_scz_bip_nointer  <- metrics_avg %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("SCZ", "BIP")) %>% 
filter(Interaction == "No-Inter") %>% 
filter(str_detect(Features, "\\+"))

metrics_scz_bip_nointer_linear  <- metrics_avg %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("SCZ", "BIP")) %>% 
filter(Interaction == "No-Inter" & Model_type == "Linear") %>% 
filter(str_detect(Features, "\\+"))



metrics_scz_bip_nointer_DL  <- metrics_avg %>%
filter(CV == "5-CV") %>% 
filter(Diagnosis %in% c("SCZ", "BIP")) %>% 
filter(Interaction == "Inter" & Model_type == "DL") %>% 
filter(str_detect(Features, "\\+"))

metrics_scz_bip_nointer_DL %>% count(Features)
metrics_scz_bip_nointer_linear %>% count(Features)
metrics_adhd_asd_mdd_inter %>% count(Features)


# metrics_multiple_inter <- bind_rows(metrics_adhd_asd_mdd_inter, metrics_scz_bip_nointer)
# metrics_multiple_inter_repl <- bind_rows(metrics_adhd_asd_mdd_inter_repl, metrics_scz_bip_nointer_repl)
metrics_multiple_inter <- bind_rows(metrics_adhd_asd_mdd_inter, metrics_scz_bip_nointer_linear, metrics_scz_bip_nointer_DL)


metrics_adhd_asd_mdd_inter %>% count(Diagnosis, Model_type, Interaction)
metrics_scz_bip_nointer_linear %>% count(Diagnosis, Model_type, Interaction)
metrics_scz_bip_nointer_DL %>% count(Diagnosis, Model_type, Interaction)
metrics_multiple_inter %>% count(Diagnosis, Model_type, Interaction)

metrics_adhd_asd_mdd_inter %>% count(Diagnosis, Model_type, Model, Features)
metrics_scz_bip_nointer_linear %>% count(Diagnosis, Model_type, Model, Features)
metrics_scz_bip_nointer_DL %>% count(Diagnosis, Model_type, Model, Features)
metrics_multiple_inter %>% count(Diagnosis, Model_type, Model, Features)


metrics_dl_vs_linear_avg  <- bind_rows(
  metrics_big,
  # metrics_big_eir,
  metrics_gln,
  # metrics_gln_logreg,
  metrics_pgs_fgrs_eir,
  metrics_pgs_fgrs_logreg,
  metrics_multiple_inter
) %>% 
filter_out_ADHD() %>% 
distinct()


```


## P-values nested models (different features) in same Integration (DL or Linear) on test_iPSYCH1 and iPSYCH2
Question: does integrating GLN+bigstatsr improve the performance of the model compared to them singularly?
Question: does adding PGS to FGRS improve the performance of the model?
Question: does adding GLN or bigstatsr to PGS+FGRS improve the performance of the model?

### test_iPSYCH1
```{r}

features_dl_vs_linear  <- c(
  "GLN", 
  "bigstatsr", 
  "GLN+bigstatsr",
  "PGS", 
  "FGRS", 
  "PGS+FGRS",
  "GLN+PGS+FGRS",
  "bigstatsr+PGS+FGRS",
  "GLN+bigstatsr+PGS+FGRS"
)

# Because of the damn ggpubr, here I need to re-compute only the p-values I wanna show on the plot, otherwise the position of the y is messed up.
to_compare <- "different_features_same_eval"
comparison_params <- set_comparison_parameters(to_compare)

nested_models_test_iPSYCH1_toplot <- metrics_dl_vs_linear_repl %>%  
filter_out_ADHD() %>% 
filter(Split_eval == "test_iPSYCH1") %>%
filter(Features %in% comparison_params$runs_to_compare_list) %>% 
filter(Integration == "Linear" | (Integration == "DL" & Model_genotype == "GLN"))


# Here I had to remove specific comparisons because they are DIagnosis specific

folder_name <- file.path(final_folder, glue("emp_pvalues_{date}_{to_compare}"))
filename  <- "auc_pvalues_nested_models_test_iPSYCH1_toplot"
# file.remove(file.path(current_dir, folder_name, filename))
if (!file.exists(file.path(folder_name, filename))) {
auc_pvalues_nested_models_test_iPSYCH1_toplot_linear <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = nested_models_test_iPSYCH1_toplot %>% mutate(Model = "Linear", Integration = "Linear", Model_type = "Linear"),
  # columns_to_group = comparison_params$columns_to_group,
  columns_to_group = c("Diagnosis", "Model", "Model_type","CV","Split_trained","Split_eval"),
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Features",
  save_plots = TRUE,
  folder_name = folder_name,
  correct_multiple = FALSE
) %>% 
filter(!(Diagnosis == "ADHD" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "ASD" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "BIP" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "MDD" & Model1 == "FGRS")) %>%
filter(!(Diagnosis == "SCZ" & Model1 == "FGRS"))  %>% 
apply_correct_multiple(., correct_multiple = TRUE); auc_pvalues_nested_models_test_iPSYCH1_toplot_linear

saveRDS(auc_pvalues_nested_models_test_iPSYCH1_toplot_linear, file.path(folder_name, filename))
} else {
auc_pvalues_nested_models_test_iPSYCH1_toplot_linear <- readRDS(file.path(folder_name, filename))
}



metrics_linear_test  <- metrics_dl_vs_linear_avg %>% 
filter(Split_eval == "test_iPSYCH1") %>% 
filter(Features %in% features_dl_vs_linear) %>%
mutate(Features = fct_relevel(Features, features_dl_vs_linear)) %>% 
filter(Model_type == "Linear" | (Model_type == "DL" & Model_genotype == "GLN"))




# ----------------------------------
# R2 p-values
# ----------------------------------
filename  <- file.path("r2_pvalues_nested_models_test_iPSYCH1_toplot")
# file.remove(file.path(current_dir, folder_name, filename))
if (!file.exists(file.path(folder_name, filename))) {
r2_pvalues_nested_models_test_iPSYCH1_toplot_linear <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = nested_models_test_iPSYCH1_toplot %>% mutate(Model = "Linear", Integration = "Linear", Model_type = "Linear"),
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Features",
  metric = "r2",
  folder_name = folder_name,
  correct_multiple = FALSE
) %>% 
filter(!(Diagnosis == "ADHD" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "ASD" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "BIP" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "MDD" & Model1 == "FGRS")) %>%
filter(!(Diagnosis == "SCZ" & Model1 == "FGRS"))  %>% 
apply_correct_multiple(., correct_multiple = TRUE); r2_pvalues_nested_models_test_iPSYCH1_toplot_linear
saveRDS(r2_pvalues_nested_models_test_iPSYCH1_toplot_linear, file.path(folder_name, filename))
} else {
r2_pvalues_nested_models_test_iPSYCH1_toplot_linear <- readRDS(file.path(folder_name, filename))
}


```

### iPSYCH2
```{r}

to_compare <- "different_features_same_eval"
comparison_params <- set_comparison_parameters(to_compare)

nested_models_iPSYCH2_toplot <- metrics_dl_vs_linear_repl %>%  
filter_out_ADHD() %>% 
filter(Split_eval == "iPSYCH2") %>%
filter(Features %in% comparison_params$runs_to_compare_list) %>% 
filter(Integration == "Linear" | (Integration == "DL" & Model_genotype == "GLN"))


folder_name <- file.path(final_folder, glue("emp_pvalues_{date}_{to_compare}"))
filename  <- "auc_pvalues_nested_models_iPSYCH2_toplot_linear"
# file.remove(file.path(current_dir, folder_name, filename))
if (!file.exists(file.path(folder_name, filename))) {
auc_pvalues_nested_models_iPSYCH2_toplot_linear <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = nested_models_iPSYCH2_toplot %>% mutate(Model = "Linear", Integration = "Linear", Model_type = "Linear"),
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Features",
  save_plots = TRUE,
  folder_name = folder_name
) %>% 
filter(!(Diagnosis == "ADHD" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "ASD" & Model1 == "PGS")) %>% 
filter(!(Diagnosis == "BIP" & Model1 == "FGRS")) %>% 
filter(!(Diagnosis == "MDD" & Model1 == "FGRS")) %>%
filter(!(Diagnosis == "SCZ" & Model1 == "FGRS"))  %>% 
apply_correct_multiple(., correct_multiple = TRUE); auc_pvalues_nested_models_iPSYCH2_toplot_linear
saveRDS(auc_pvalues_nested_models_iPSYCH2_toplot_linear, file.path(folder_name, filename))
} else {
auc_pvalues_nested_models_iPSYCH2_toplot_linear <- readRDS(file.path(folder_name, filename))
}


metrics_linear_iPSYCH2  <- metrics_dl_vs_linear_avg %>% 
filter(Split_eval == "iPSYCH2") %>% 
filter(Features %in% features_dl_vs_linear) %>%
mutate(Features = fct_relevel(Features, features_dl_vs_linear)) %>% 
filter(Model_type == "Linear" | (Model_type == "DL" & Model_genotype == "GLN"))




# ----------------------------------
# R2 p-values
# ----------------------------------
filename  <- "r2_pvalues_nested_models_iPSYCH2_toplot_linear"
# file.remove(file.path(current_dir, folder_name, filename))
if (!file.exists(file.path(folder_name, filename))) {
r2_pvalues_nested_models_iPSYCH2_toplot_linear <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = nested_models_iPSYCH2_toplot %>% mutate(Model = "Linear", Integration = "Linear", Model_type = "Linear") ,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Features",
  metric = "r2",
  folder_name = folder_name
); r2_pvalues_nested_models_iPSYCH2_toplot_linear
saveRDS(r2_pvalues_nested_models_iPSYCH2_toplot_linear, file.path(folder_name, filename))
} else {
r2_pvalues_nested_models_iPSYCH2_toplot_linear <- readRDS(file.path(folder_name, filename))
}



```

## P-values same features DL vs Linear test_iPSYCH1
Question: is the performance of the DL integration on the models GLN+bigstatsr, GLN+PGS+FGRS, GLN+bigstatsr+PGS+FGRS higher than the performance of the Linear integration on the same models?
```{r}
to_compare <- "DL_vs_Linear"

comparison_params <- set_comparison_parameters(to_compare)

trial_res_tabs_DL_vs_Linear <- metrics_dl_vs_linear_repl %>% 
filter_out_ADHD() %>% 
# select(GLN, Interaction) %>% 
# filter_out_ADHD() %>% 
# filter(Diagnosis %in% c("ASD")) %>%
  filter(Split_eval == "test_iPSYCH1") %>%
  filter(Integration %in% comparison_params$runs_to_compare_list) %>% 
  filter(Features %in% c("GLN+bigstatsr", "GLN+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS"))

folder_name <- file.path(final_folder, glue("emp_pvalues_{date}_{to_compare}"))
filename  <- "auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1"
# file.remove(file.path(current_dir, folder_name, filename))
if (!file.exists(file.path(folder_name, filename))) {
auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1 <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = trial_res_tabs_DL_vs_Linear,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Integration",
  folder_name = folder_name
); auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1
saveRDS(auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1, file.path(folder_name, filename))
} else {
auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1 <- readRDS(file.path(folder_name, filename))
}

filename  <- "r2_pvalues_tabs_DL_vs_Linear"
# file.remove(file.path(current_dir, folder_name, filename))
if (!file.exists(file.path(folder_name, filename))) {
r2_pvalues_tabs_DL_vs_Linear <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = trial_res_tabs_DL_vs_Linear ,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Integration",
  metric = "r2",
  folder_name = folder_name
); r2_pvalues_tabs_DL_vs_Linear
saveRDS(r2_pvalues_tabs_DL_vs_Linear, file.path(folder_name, filename))
} else {
r2_pvalues_tabs_DL_vs_Linear <- readRDS(file.path(folder_name, filename))
}


```


## Plot AUC tabs Linear test_iPSYCH1

```{r}


#-
auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear <- process_statistical_tests(
  metrics_df = metrics_linear_test %>% metrics_to_wide(metrics_auroc),
  emp_pvalues_df = auc_pvalues_nested_models_test_iPSYCH1_toplot_linear,
  metric_name = "roc_auc",
  x_col = "Features", # HERE IT'S IMPORTANT THAT IT'S THE SAAAAAME
  fill_col = "Features",
  columns_to_group = c("Diagnosis", "Split_eval"),
  nested_models = TRUE,
  step_increase = 0.02,
  keep_group_model_order = TRUE,
  print = FALSE,
  runs_to_compare = NULL,
  fun_to_apply = "mean_se"
) %>% 
select(-contains("sig")) %>% 
rename(p.adj = p_value_fdr_all) %>% 
add_significance("p.adj") %>% 
mutate(p_fdr = ifelse(p.adj.signif == "ns", "ns", round(p.adj, 3)))

saveRDS(auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear, file.path(final_folder, "auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
saveRDS(metrics_linear_test, file.path(final_folder, "metrics_linear_test.rds"))

#-
auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear  <- readRDS(file.path(final_folder, "auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
metrics_linear_test  <- readRDS(file.path(final_folder, "metrics_linear_test.rds"))


p_auc_tabs_linear_test  <- plot_barplots_ggpubr(dataframe = metrics_linear_test,
  x_column = "Features",
  fill_column = "Features",
  fill_label = "Features",
  title = "",
  x_title = "",
  metrics = metrics_auroc,
  y_title = "AUROC",
  # metrics = metrics_auroc,
  # y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Diagnosis",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right", # Remove legend by setting position to "none"
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  # pvalue_size = 9,
  fill_colors = features_colors_palette,
  # padding = -0.001
  padding = 0,
  bar_width = 0.9,
  theme = my_theme
  # expand = c(0, 0)
) + 
# add horizontal dashed line at 0.5
geom_hline(yintercept = 0.5, linetype = "dashed", color = "black")+
# theme(aspect.ratio = 1)+
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.5)) +
# theme(axis.ticks.x = element_line()) +
theme(strip.background = element_blank()) +
# rotate 90 degrees
theme(axis.text.x = element_text(angle = 90, hjust = 1))+
scale_y_continuous(breaks = seq(0.5, 0.77, by = 0.1), limits = c(0, 0.77), expand = c(0, 0))+
coord_cartesian(ylim = c(0.495, 0.74))+
scale_y_continuous(breaks = c(seq(0.5, 1, by = 0.05)))+
stat_pvalue_manual(
auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear %>% 
mutate(y.position = 
case_when(
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr" ~ 0.65,
  Diagnosis == "ADHD" & group2 == "PGS+FGRS" ~ 0.66,
  Diagnosis == "ADHD" & group2 == "GLN+PGS+FGRS" ~ 0.705,
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.72,

  Diagnosis == "ASD" & group2 == "GLN+bigstatsr" ~ 0.59,
  Diagnosis == "ASD" & group2 == "PGS+FGRS" ~ 0.61,
  Diagnosis == "ASD" & group2 == "GLN+PGS+FGRS" ~ 0.63,
  Diagnosis == "ASD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.645,

  Diagnosis == "BIP" & group2 == "GLN+bigstatsr" ~ 0.60,
  Diagnosis == "BIP" & group2 == "PGS+FGRS" ~ 0.71,
  Diagnosis == "BIP" & group2 == "GLN+PGS+FGRS" ~ 0.72,
  Diagnosis == "BIP" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.73,

  Diagnosis == "MDD" & group2 == "GLN+bigstatsr" ~ 0.61,
  Diagnosis == "MDD" & group2 == "PGS+FGRS" ~ 0.66,
  Diagnosis == "MDD" & group2 == "GLN+PGS+FGRS" ~ 0.67,
  Diagnosis == "MDD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.68,

  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr" ~ 0.62,
  Diagnosis == "SCZ" & group2 == "PGS+FGRS" ~ 0.675,
  Diagnosis == "SCZ" & group2 == "GLN+PGS+FGRS" ~ 0.68,
  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.69,
  TRUE ~ y.position
)
),
label = "p_fdr", 
tip.length = 0.01, 
size = pvalue_size_fig3)+
# remove legend
theme(legend.position = "none"); p_auc_tabs_linear_test
# hide.ns = TRUE);p_auc_tabs_linear_test
# hide.ns = "p.adj");p_auc_tabs_linear_test

ggsave(file.path(figs_path, "auc_tabs_linear_testiPSYCH1.pdf"), print(p_auc_tabs_linear_test), width = 70, height = 25, units = "cm")

# TODO: 16/02/25 16:13 this works



```
## Plot R2 tabs Linear test_iPSYCH1
```{r}

#-
r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear <- process_statistical_tests(
  metrics_df = metrics_linear_test %>% metrics_to_wide(metrics_r2),
  emp_pvalues_df = r2_pvalues_nested_models_test_iPSYCH1_toplot_linear,
  metric_name = "r2",
  x_col = "Features", # HERE IT'S IMPORTANT THAT IT'S THE SAAAAAME
  fill_col = "Features",
  columns_to_group = c("Diagnosis", "Split_eval"),
  nested_models = TRUE,
  step_increase = 0.12,
  keep_group_model_order = TRUE,
  print = TRUE,
  runs_to_compare = NULL,
  fun_to_apply = "mean"
) %>% 
select(-contains("sig")) %>% 
rename(p.adj = p_value_fdr_all) %>% 
add_significance("p.adj") %>% 
mutate(p_fdr = ifelse(p.adj.signif == "ns", "ns", round(p.adj, 3)))


saveRDS(r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear, file.path(final_folder, "r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
saveRDS(metrics_linear_test, file.path(final_folder, "metrics_linear_test.rds"))


#-
r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear  <- readRDS(file.path(final_folder, "r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
metrics_linear_test  <- readRDS(file.path(final_folder, "metrics_linear_test.rds"))


metrics_linear_test %>% 
filter(!metric %in% c("prevalence","m_case")) %>% 
mutate(metric_type = case_when(
  str_detect(metric, "r2_lin_liab_Florian") ~ "r2_lin_liab_Florian", 
  str_detect(metric, "r2_lin_liab_ABC") ~ "r2_lin_liab_ABC", 
  str_detect(metric, "r2_lin") ~ "r2_lin",
  str_detect(metric, "r2_nag_liab") ~ "r2_nag_liab_ABC",
  str_detect(metric, "r2_nag") ~ "r2_nag",
  str_detect(metric, "auc") ~ "auc",
  TRUE ~ metric
)) %>%
mutate(metric = str_extract(metric, "^[^_]+")) %>%
pivot_wider(names_from = metric, values_from = estimate) %>% 
mutate(metric_type = fct_relevel(metric_type, c("auc", "r2_lin", "r2_lin_liab_Florian", "r2_lin_liab_ABC", "r2_nag", "r2_nag_liab_ABC"))) %>% 
# filter(!metric_type %in% c("auc")) %>% 
ggbarplot(
  x = "Features",
  y = "avg", 
  fill = "Features",
  facet.by = c("metric_type", "Diagnosis"),
  scales = "free",
  color = "black",
  width = 0.9 # Set width=1 to remove space between bars
)+
geom_errorbar(
  aes(
    ymin = lower,
    ymax = upper,
    group = interaction(Features, Features)
  ),
  width = 0.2,
  position = position_dodge(width = 0.9) # Match width=1 from ggbarplot
)+
scale_fill_manual(values = features_colors_palette)+
my_theme+
theme(axis.text.x = element_blank())

#-
r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear  <- readRDS(file.path(final_folder, "r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
metrics_linear_test  <- readRDS(file.path(final_folder, "metrics_linear_test.rds"))

p_r2_tabs_linear_test  <- plot_barplots_ggpubr(dataframe = metrics_linear_test,
  x_column = "Features",
  fill_column = "Features",
  fill_label = "Features",
  title = "",
  x_title = "",
  metrics = metrics_r2,
  y_title = "R2",
  # metrics = metrics_auroc,
  # y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Diagnosis",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right",
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  fill_colors = features_colors_palette,
  # padding = -0.001
  padding = 0,
  bar_width = 0.9,
  theme = my_theme
  # expand = c(0, 0)
) + 
# add horizontal dashed line at 0.5
# theme(aspect.ratio = 1)+
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.5)) +
theme(strip.background = element_blank()) +
theme(axis.text.x = element_text(angle = 90, hjust = 1))+
stat_pvalue_manual(
r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear %>% 
mutate(y.position = 
case_when(
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr" ~ 0.06,
  Diagnosis == "ADHD" & group2 == "PGS+FGRS" ~ 0.09,
  Diagnosis == "ADHD" & group2 == "GLN+PGS+FGRS" ~ 0.105,
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.115,

  Diagnosis == "ASD" & group2 == "GLN+bigstatsr" ~ 0.02,
  Diagnosis == "ASD" & group2 == "PGS+FGRS" ~ 0.04,
  Diagnosis == "ASD" & group2 == "GLN+PGS+FGRS" ~ 0.05,
  Diagnosis == "ASD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.06,

  Diagnosis == "BIP" & group2 == "GLN+bigstatsr" ~ 0.02,
  Diagnosis == "BIP" & group2 == "PGS+FGRS" ~ 0.225,
  Diagnosis == "BIP" & group2 == "GLN+PGS+FGRS" ~ 0.235 ,
  Diagnosis == "BIP" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.245,

  Diagnosis == "MDD" & group2 == "GLN+bigstatsr" ~ 0.03,
  Diagnosis == "MDD" & group2 == "PGS+FGRS" ~ 0.095,
  Diagnosis == "MDD" & group2 == "GLN+PGS+FGRS" ~ 0.105,
  Diagnosis == "MDD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.115,

  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr" ~ 0.025,
  Diagnosis == "SCZ" & group2 == "PGS+FGRS" ~ 0.09,
  Diagnosis == "SCZ" & group2 == "GLN+PGS+FGRS" ~ 0.105,
  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.115,
  TRUE ~ y.position
)
),
label = "p_fdr", 
tip.length = 0.01, 
size = pvalue_size);p_r2_tabs_linear_test



r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear %>% 
select(Diagnosis, group1, group2, p_value, y.position)

ggsave(file.path(figs_path, "r2_tabs_linear_test_pvalues.pdf"), print(p_r2_tabs_linear_test), width = 70, height = 25, units = "cm")


```


## P-values same features DL vs Linear iPSYCH2
Question: is the performance of the DL integration on the models GLN+bigstatsr, GLN+PGS+FGRS, GLN+bigstatsr+PGS+FGRS higher than the performance of the Linear integration on the same models?
```{r}
to_compare <- "DL_vs_Linear"

comparison_params <- set_comparison_parameters(to_compare)

trial_res_tabs_DL_vs_Linear_iPSYCH2 <- metrics_dl_vs_linear_repl %>% 
filter_out_ADHD() %>% 
# select(GLN, Interaction) %>% 
# filter_out_ADHD() %>% 
# filter(Diagnosis %in% c("ASD")) %>%
  filter(Split_eval == "iPSYCH2") %>%
  filter(Integration %in% comparison_params$runs_to_compare_list) %>% 
  filter(Features %in% c("GLN+bigstatsr", "GLN+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS"))

folder_name <- file.path(final_folder, glue("emp_pvalues_{date}_{to_compare}"))
filename  <- "auc_pvalues_tabs_DL_vs_Linear_iPSYCH2"
# remove file
# if (file.exists(file.path(current_dir, folder_name, filename))) {file.remove(file.path(current_dir, folder_name, filename))}


if (!file.exists(file.path(folder_name, filename))) {
auc_pvalues_tabs_DL_vs_Linear_iPSYCH2 <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = trial_res_tabs_DL_vs_Linear_iPSYCH2,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Integration",
  folder_name = folder_name
); auc_pvalues_tabs_DL_vs_Linear_iPSYCH2
saveRDS(auc_pvalues_tabs_DL_vs_Linear_iPSYCH2, file.path(folder_name, filename))
} else {
auc_pvalues_tabs_DL_vs_Linear_iPSYCH2 <- readRDS(file.path(folder_name, filename))
}

filename  <- "r2_pvalues_tabs_DL_vs_Linear_iPSYCH2"
if (!file.exists(file.path(folder_name, filename))) {
r2_pvalues_tabs_DL_vs_Linear_iPSYCH2 <- compute_pvalues_comparisons(
  to_compare = to_compare,
  data = trial_res_tabs_DL_vs_Linear_iPSYCH2 ,
  columns_to_group = comparison_params$columns_to_group,
  runs_to_compare = comparison_params$runs_to_compare,
  filter_column = "Integration",
  metric = "r2",
  folder_name = folder_name
); r2_pvalues_tabs_DL_vs_Linear_iPSYCH2
saveRDS(r2_pvalues_tabs_DL_vs_Linear_iPSYCH2, file.path(folder_name, filename))
} else {
r2_pvalues_tabs_DL_vs_Linear_iPSYCH2 <- readRDS(file.path(folder_name, filename))
}

```


## Plot AUC tabs Linear iPSYCH2
```{r}

#-
auc_pvalues_formatted_nested_models_iPSYCH2_toplot_linear <- process_statistical_tests(
  metrics_df = metrics_linear_iPSYCH2 %>% metrics_to_wide(metrics_auroc),
  emp_pvalues_df = auc_pvalues_nested_models_iPSYCH2_toplot_linear,
  metric_name = "roc_auc",
  x_col = "Features", # HERE IT'S IMPORTANT THAT IT'S THE SAAAAAME
  fill_col = "Features",
  columns_to_group = c("Diagnosis", "Split_eval"),
  nested_models = TRUE,
  step_increase = 0.02,
  keep_group_model_order = TRUE,
  print = FALSE,
  runs_to_compare = NULL,
  fun_to_apply = "mean_se"
) %>%
select(-contains("sig")) %>% 
rename(p.adj = p_value_fdr_all) %>% 
add_significance("p.adj") %>% 
mutate(p_fdr = ifelse(p.adj.signif == "ns", "ns", round(p.adj, 3)))


saveRDS(auc_pvalues_formatted_nested_models_iPSYCH2_toplot_linear, file.path(final_folder, "auc_pvalues_formatted_nested_models_iPSYCH2_toplot_linear.rds"))
saveRDS(metrics_linear_iPSYCH2, file.path(final_folder, "metrics_linear_iPSYCH2.rds"))

p_auc_tabs_linear_iPSYCH2  <- plot_barplots_ggpubr(dataframe = metrics_linear_iPSYCH2,
  x_column = "Features",
  fill_column = "Features",
  fill_label = "Features",
  title = "",
  x_title = "",
  metrics = metrics_auroc,
  y_title = "AUROC",
  # metrics = metrics_auroc,
  # y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Diagnosis",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right",
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  fill_colors = features_colors_palette,
  # padding = -0.001
  padding = 0,
  bar_width = 0.9,
  theme = my_theme
  # expand = c(0, 0)
) + 
# add horizontal dashed line at 0.5
geom_hline(yintercept = 0.5, linetype = "dashed", color = "black")+
# theme(aspect.ratio = 1)+
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.5)) +
# theme(axis.ticks.x = element_line()) +
theme(strip.background = element_blank()) +
# rotate 90 degrees
theme(axis.text.x = element_text(angle = 90, hjust = 1))+
scale_y_continuous(breaks = seq(0.5, 0.77, by = 0.1), limits = c(0, 0.77), expand = c(0, 0))+
coord_cartesian(ylim = c(0.495, 0.77))+
scale_y_continuous(breaks = c(seq(0.5, 1, by = 0.05)))+
stat_pvalue_manual(auc_pvalues_formatted_nested_models_iPSYCH2_toplot_linear %>% 
mutate(y.position = 
case_when(
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr" ~ 0.62,
  Diagnosis == "ADHD" & group2 == "PGS+FGRS" ~ 0.63,
  Diagnosis == "ADHD" & group2 == "GLN+PGS+FGRS" ~ 0.68,
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.69,

  Diagnosis == "ASD" & group2 == "GLN+bigstatsr" ~ 0.58,
  Diagnosis == "ASD" & group2 == "PGS+FGRS" ~ 0.60,
  Diagnosis == "ASD" & group2 == "GLN+PGS+FGRS" ~ 0.62,
  Diagnosis == "ASD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.635,

  Diagnosis == "BIP" & group2 == "GLN+bigstatsr" ~ 0.55,
  Diagnosis == "BIP" & group2 == "PGS+FGRS" ~ 0.675,
  Diagnosis == "BIP" & group2 == "GLN+PGS+FGRS" ~ 0.685,
  Diagnosis == "BIP" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.695,

  Diagnosis == "MDD" & group2 == "GLN+bigstatsr" ~ 0.56,
  Diagnosis == "MDD" & group2 == "PGS+FGRS" ~ 0.63,
  Diagnosis == "MDD" & group2 == "GLN+PGS+FGRS" ~ 0.645,
  Diagnosis == "MDD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.655,

  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr" ~ 0.565,
  Diagnosis == "SCZ" & group2 == "PGS+FGRS" ~ 0.645,
  Diagnosis == "SCZ" & group2 == "GLN+PGS+FGRS" ~ 0.65,
  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.66,
  TRUE ~ y.position
)
),
label = "p_fdr", 
tip.length = 0.01, 
size = pvalue_size,
hide.ns = FALSE);p_auc_tabs_linear_iPSYCH2
# hide.ns = "p.adj");prova

ggsave(file.path(figs_path, "auc_tabs_linear_iPSYCH2_pvalues.pdf"), p_auc_tabs_linear_iPSYCH2, width = 70, height = 25, units = "cm")




```


## Plot R2 tabs Linear iPSYCH2
```{r}


#-
r2_pvalues_formatted_nested_models_iPSYCH2_toplot_linear <- process_statistical_tests(
  metrics_df = metrics_linear_iPSYCH2 %>% metrics_to_wide(metrics_r2),
  emp_pvalues_df = r2_pvalues_nested_models_iPSYCH2_toplot_linear,
  metric_name = "r2",
  x_col = "Features", # HERE IT'S IMPORTANT THAT IT'S THE SAAAAAME
  fill_col = "Features",
  columns_to_group = c("Diagnosis", "Split_eval"),
  nested_models = TRUE,
  step_increase = 0.12,
  keep_group_model_order = TRUE,
  print = FALSE,
  runs_to_compare = NULL,
  fun_to_apply = "mean_se"
) %>% 
select(-contains("sig")) %>% 
rename(p.adj = p_value_fdr_all) %>% 
add_significance("p.adj") %>% 
mutate(p_fdr = ifelse(p.adj.signif == "ns", "ns", round(p.adj, 3)))

saveRDS(metrics_linear_iPSYCH2, file.path(final_folder, "metrics_linear_iPSYCH2.rds"))

saveRDS(r2_pvalues_formatted_nested_models_iPSYCH2_toplot_linear, file.path(final_folder, "r2_pvalues_formatted_nested_models_iPSYCH2_toplot_linear.rds"))

p_r2_tabs_linear_iPSYCH2  <- plot_barplots_ggpubr(dataframe = metrics_linear_iPSYCH2,
  x_column = "Features",
  fill_column = "Features",
  fill_label = "Features",
  title = "",
  x_title = "",
  metrics = metrics_r2,
  y_title = "R2",
  # metrics = metrics_auroc,
  # y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Diagnosis",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right",
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  fill_colors = features_colors_palette,
  # padding = -0.001
  padding = 0,
  bar_width = 0.9,
  theme = my_theme
  # expand = c(0, 0)
) + 
# add horizontal dashed line at 0.5
# theme(aspect.ratio = 1)+
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.5)) +
theme(strip.background = element_blank()) +
theme(axis.text.x = element_text(angle = 90, hjust = 1))+
stat_pvalue_manual(r2_pvalues_formatted_nested_models_iPSYCH2_toplot_linear %>% 
mutate(y.position = 
case_when(
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr" ~ 0.04,
  Diagnosis == "ADHD" & group2 == "PGS+FGRS" ~ 0.085,
  Diagnosis == "ADHD" & group2 == "GLN+PGS+FGRS" ~ 0.10,
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.105,

  Diagnosis == "ASD" & group2 == "GLN+bigstatsr" ~ 0.015,
  Diagnosis == "ASD" & group2 == "PGS+FGRS" ~ 0.04,
  Diagnosis == "ASD" & group2 == "GLN+PGS+FGRS" ~ 0.045,
  Diagnosis == "ASD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.05,

  Diagnosis == "BIP" & group2 == "GLN+bigstatsr" ~ 0.01,
  Diagnosis == "BIP" & group2 == "PGS+FGRS" ~ 0.10,
  Diagnosis == "BIP" & group2 == "GLN+PGS+FGRS" ~ 0.105 ,
  Diagnosis == "BIP" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.11,

  Diagnosis == "MDD" & group2 == "GLN+bigstatsr" ~ 0.015,
  Diagnosis == "MDD" & group2 == "PGS+FGRS" ~ 0.06,
  Diagnosis == "MDD" & group2 == "GLN+PGS+FGRS" ~ 0.065,
  Diagnosis == "MDD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.07,

  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr" ~ 0.01,
  Diagnosis == "SCZ" & group2 == "PGS+FGRS" ~ 0.055,
  Diagnosis == "SCZ" & group2 == "GLN+PGS+FGRS" ~ 0.06,
  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.065,
  TRUE ~ y.position
)
),
label = "p_fdr", 
tip.length = 0.01, 
size = pvalue_size,
hide.ns = FALSE);p_r2_tabs_linear_iPSYCH2

ggsave(file.path(figs_path, "r2_tabs_linear_iPSYCH2_pvalues.pdf"), print(p_r2_tabs_linear_iPSYCH2), width = 70, height = 25, units = "cm")

```


## AUC Performance drop tabs DL vs Linear test iPSYCH1
```{r}

#-
features_drop_perf_tab  <- c("GLN+bigstatsr", "GLN+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS")

perf_drop_test_auc_tabs <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter(Split_eval == "test_iPSYCH1") %>% select(-any_of(c("Interaction", "Integration"))),
  features_to_compare = features_drop_perf_tab,
  metrics_cols = metrics_auroc, 
  cols_to_remove = c("CV", "Label" , "Model_genotype", "Split_eval", "Split_trained", "Model"),
  category_col = "Model_type",
  split1 = "Linear",
  split2 = "DL"
) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab))


# ----------------------------------
# Within-model pvalues DL-Linear
# ----------------------------------

# I had calculated the pvalues for the within-models DL_vs_linear here auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1, so I'll just left_join it

perf_drop_test_auc_tabs  <- perf_drop_test_auc_tabs %>% 
left_join(auc_pvalues_tabs_DL_vs_Linear_testiPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab)) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))


saveRDS(perf_drop_test_auc_tabs, file.path(final_folder, "perf_drop_test_auc_tabs.rds"))

#-
perf_drop_test_auc_tabs <- readRDS(file.path(final_folder, "perf_drop_test_auc_tabs.rds"))

bar_width  <- 0.7
pvalue_size_fig3 <- 7
p_perf_drop_tabs_test_DL_linear_auc <- ggbarplot(perf_drop_test_auc_tabs,
  x = "Features",
  y = "diff_avg",
  fill = "Features",
  position = position_dodge(width = bar_width, preserve = "single"),
  width = bar_width,
  add = "",
  color = "black"
) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * AUROC ~ .(suffix_delta_model)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Features
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.006,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = +1, # Adjust vertical spacing
    size = pvalue_size_fig3
  ) +
  scale_y_reverse() +
  my_theme +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  facet_grid(~Diagnosis, scales = "free_x"); p_perf_drop_tabs_test_DL_linear_auc


```

## R2 Performance drop tabs DL vs Linear test iPSYCH1
```{r}


perf_drop_test_r2_tabs <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter(Split_eval == "test_iPSYCH1")  %>% select(-any_of(c("Interaction", "Integration"))),
  features_to_compare = features_drop_perf_tab,
  metrics_cols = metrics_r2, 
  cols_to_remove = c("CV", "Label" , "Model_genotype", "Split_eval", "Split_trained", "Model"),
  category_col = "Model_type",
  split1 = "Linear",
  split2 = "DL"
) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab))


# I had calculated the pvalues for the within-models DL_vs_linear here r2_pvalues_tabs_DL_vs_Linear, so I'll just left_join it

perf_drop_test_r2_tabs  <- perf_drop_test_r2_tabs %>% 
left_join(r2_pvalues_tabs_DL_vs_Linear, by = c("Diagnosis", "Features")) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab)) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))


saveRDS(perf_drop_test_r2_tabs, file.path(final_folder, "perf_drop_test_r2_tabs.rds"))

#-

perf_drop_test_r2_tabs <- readRDS(file.path(final_folder, "perf_drop_test_r2_tabs.rds"))
bar_width  <- 0.7
p_perf_drop_tabs_test_DL_linear_r2 <- ggbarplot(perf_drop_test_r2_tabs,
  x = "Features",
  y = "diff_avg",
  fill = "Features",
  position = position_dodge(width = bar_width, preserve = "single"),
  width = bar_width,
  add = "",
  color = "black"
) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * R^2 * .(suffix_delta_model)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Features
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.006,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5, # Adjust vertical spacing
    size = pvalue_size
  ) +
  scale_y_reverse() +
  my_theme +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  facet_grid(~Diagnosis, scales = "free_x");p_perf_drop_tabs_test_DL_linear_r2




```


## AUC Performance drop tabs DL vs Linear iPSYCH2
```{r}

#-
features_drop_perf_tab  <- c("GLN+bigstatsr", "GLN+PGS+FGRS", "GLN+bigstatsr+PGS+FGRS")

perf_drop_iPSYCH2_auc_tabs <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter(Split_eval == "iPSYCH2") %>% select(-any_of(c("Interaction", "Integration"))),
  features_to_compare = features_drop_perf_tab,
  metrics_cols = metrics_auroc, 
  cols_to_remove = c("CV", "Label" , "Model_genotype", "Split_eval", "Split_trained", "Model"),
  category_col = "Model_type",
  split1 = "Linear",
  split2 = "DL"
) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab))


# ----------------------------------
# Within-model pvalues DL-Linear
# ----------------------------------

# I had calculated the pvalues for the within-models DL_vs_linear here r2_pvalues_tabs_DL_vs_Linear_iPSYCH2, so I'll just left_join it

perf_drop_iPSYCH2_auc_tabs  <- perf_drop_iPSYCH2_auc_tabs %>% 
left_join(auc_pvalues_tabs_DL_vs_Linear_iPSYCH2, by = c("Diagnosis", "Features")) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab)) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

saveRDS(perf_drop_iPSYCH2_auc_tabs, file.path(final_folder, "perf_drop_iPSYCH2_auc_tabs.rds"))

bar_width  <- 0.7
p_perf_drop_tabs_iPSYCH2_DL_linear_auc <- ggbarplot(perf_drop_iPSYCH2_auc_tabs,
  x = "Features",
  y = "diff_avg",
  fill = "Features",
  position = position_dodge(width = bar_width, preserve = "single"),
  width = bar_width,
  add = "",
  color = "black"
) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * AUROC * .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Features
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.003,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5, # Adjust vertical spacing
    size = pvalue_size
  ) +
  scale_y_reverse() +
  my_theme +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  facet_grid(~Diagnosis, scales = "free_x");p_perf_drop_tabs_iPSYCH2_DL_linear_auc

ggpar(p_perf_drop_tabs_iPSYCH2_DL_linear_auc, ylim = c(+0.05, -0.05))
# Add p-values


# fig3_auc  <- (p_auc_tabs_linear_iPSYCH2 + theme(legend.position = "bottom")) / (p_perf_drop_tabs_iPSYCH2_DL_linear_auc + theme(legend.position = "bottom") + plot_spacer())

# fig3_auc  <- (p_auc_tabs_linear_iPSYCH2 + theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1))) / (p_perf_drop_tabs_iPSYCH2_DL_linear_auc + theme(legend.position = "bottom") + plot_spacer())

p_tabs_test_auc_iPSYCH2  <- (p_auc_tabs_linear_iPSYCH2 + theme(legend.position = "none")) / (p_perf_drop_tabs_iPSYCH2_DL_linear_auc + theme(legend.position = "none") + plot_spacer()) + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(face = "bold"))

ggsave(glue("plots_paper_final/figs_{date}/perf_drop_tabs_iPSYCH2_DL_linear_auc.pdf"), p_perf_drop_tabs_iPSYCH2_DL_linear_auc, width = 50, height = 30, units = "cm")

width  <- 90
height <- 60
ggsave(glue("plots_paper_final/figs_{date}/auc_linear_iPSYCH2_and_auc_drop_tabs_DL_linear_iPSYCH2_{width}x{height}.pdf"), p_tabs_test_auc_iPSYCH2, width = width, height = height, units = "cm")

```

## R2 Performance drop tabs DL vs Linear iPSYCH2
```{r}

#-
perf_drop_iPSYCH2_r2_tabs <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter(Split_eval == "iPSYCH2") %>% select(-any_of(c("Interaction", "Integration"))),
  features_to_compare = features_drop_perf_tab,
  metrics_cols = metrics_r2, 
  cols_to_remove = c("CV", "Label" , "Model_genotype", "Split_eval", "Split_trained", "Model"),
  category_col = "Model_type",
  split1 = "Linear",
  split2 = "DL"
) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab))


# I had calculated the pvalues for the within-models DL_vs_linear here r2_pvalues_tabs_DL_vs_Linear, so I'll just left_join it

perf_drop_iPSYCH2_r2_tabs  <- perf_drop_iPSYCH2_r2_tabs %>% 
left_join(r2_pvalues_tabs_DL_vs_Linear_iPSYCH2, by = c("Diagnosis", "Features")) %>% 
mutate(Features = factor(Features, levels = features_drop_perf_tab)) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

saveRDS(perf_drop_iPSYCH2_r2_tabs, file.path(final_folder, "perf_drop_iPSYCH2_r2_tabs.rds"))

bar_width  <- 0.7
p_perf_drop_tabs_iPSYCH2_DL_linear_r2 <- ggbarplot(perf_drop_iPSYCH2_r2_tabs,
  x = "Features",
  y = "diff_avg",
  fill = "Features",
  position = position_dodge(width = bar_width, preserve = "single"),
  width = bar_width,
  add = "",
  color = "black"
) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * R^2 * .(suffix_delta_model)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Features
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.003,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5, # Adjust vertical spacing
    size = pvalue_size_fig3
  ) +
  scale_y_reverse() +
  my_theme +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  facet_grid(~Diagnosis, scales = "free_x");p_perf_drop_tabs_iPSYCH2_DL_linear_r2

ggpar(p_perf_drop_tabs_iPSYCH2_DL_linear_r2, ylim = c(+0.05, -0.05))

ggsave(file.path(figs_path, "perf_drop_tabs_iPSYCH2_DL_linear_r2.pdf"), p_perf_drop_tabs_iPSYCH2_DL_linear_r2, width = 50, height = 30, units = "cm")

p_tabs_r2_iPSYCH2  <- (p_r2_tabs_linear_iPSYCH2 + theme(legend.position = "none")) / (p_perf_drop_tabs_iPSYCH2_DL_linear_r2 + theme(legend.position = "none") + plot_spacer()) + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(face = "bold"));p_tabs_r2_iPSYCH2

width  <- 90
height <- 60
ggsave(glue("plots_paper_final/figs_{date}/r2_linear_iPSYCH2_and_r2_drop_tabs_DL_linear_iPSYCH2_{width}x{height}.pdf"), p_tabs_r2_iPSYCH2, width = width, height = height, units = "cm")

```

## AUC Linear perf drop iPSYCH2 vs test_iPSYCH1 all features
Here I want all the features but showing their drop in performance of their linear integration in iPSYCH2 vs test_iPSYCH1
```{r}

#-
features  <- c(
  "GLN",
  "bigstatsr",
  "GLN+bigstatsr",
  "PGS",
  "FGRS",
  "PGS+FGRS",
  "GLN+PGS+FGRS",
  "bigstatsr+PGS+FGRS",
  "GLN+bigstatsr+PGS+FGRS"
)
metrics_avg %>% count(Features)
# AUROC
# Calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the average values
auc_drop_all_linear_iPSYCH2_vs_test_iPSYCH1 <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter_out_ADHD() %>% filter(Split_eval != "valid") %>% filter(Model_type == "Linear" | (Model_type == "DL" & Model_genotype == "GLN")) %>% filter(CV == "5-CV"),
  # metrics_df = metrics_dl_vs_linear_avg %>% filter_out_ADHD() %>% filter(Model_type == "Linear") %>% filter(CV == "5-CV"),
  features_to_compare = features,
  cols_to_remove = c("CV", "Label", "Model_type", "Split_trained"),
  category_col = "Split_eval",
  metrics_cols = metrics_auroc,
  split1 = "test_iPSYCH1",
  split2 = "iPSYCH2"
) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))));auc_drop_all_linear_iPSYCH2_vs_test_iPSYCH1


# Format the dataframe to calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the bootstrap replicates
column_replicates <- "auc_replicates"
auc_formatted_drop_all_linear_iPSYCH2  <- metrics_dl_vs_linear_repl %>%          
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(CV == "5-CV") %>%
          filter(Model_type == "Linear" | (Model_type == "DL" & Model_genotype == "GLN")) %>% 
          # filter(Features %in% runs_to_compare[[1]]) %>%
          # mutate(Features = fct_relevel(Features, runs_to_compare[[1]])) %>% 
          mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
          select(Diagnosis, Features, Split_eval, Split_trained, CV, all_of(column_replicates)) %>% 
          pivot_wider(names_from = Split_eval, values_from = all_of(column_replicates)) %>% 
          mutate(!!sym(column_replicates) := map2(iPSYCH2, test_iPSYCH1, ~ .x - .y))


# auc_pvalues_same_feature_same_model_different_eval %>% filter(Features == "GLN+bigstatsr") 

# ----------------------------------
# Within-model pvalues
# ----------------------------------
# Get the real p-values for the GLN and bigstatsr models
auc_pvalues_all_linear_iPSYCH2_vs_iPSYCH1 <- auc_pvalues_same_feature_same_model_different_eval %>% 
filter_out_ADHD() %>% 
filter(Model_type == "Linear" | (Model_type == "DL" & Features == "GLN"))

# auc_pvalues_all_linear_iPSYCH2_vs_iPSYCH1 %>% filter(Features == "GLN+bigstatsr") # It's there, Linear


# Format the dataframe to be compatible with geom_text
auc_df_all_linear_iPSYCH2_vs_iPSYCH1  <- auc_drop_all_linear_iPSYCH2_vs_test_iPSYCH1 %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
left_join(auc_pvalues_all_linear_iPSYCH2_vs_iPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(Features = fct_relevel(Features, features)) %>%  
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

# create directory
saveRDS(auc_df_all_linear_iPSYCH2_vs_iPSYCH1, file.path(final_folder, "auc_df_all_linear_iPSYCH2_vs_iPSYCH1.rds"))

#-
# Barplot
x_column <- "Diagnosis"
y_column <- "diff_avg"
fill_column <- "Features"
metric_name <- "diff"
bar_width  <- 0.9
y_min  <- -0.05
y_max  <- 0.12

auc_df_all_linear_iPSYCH2_vs_iPSYCH1 %>% count(Features)

p_auc_perf_drop_all_linear_iPSYCH2  <- ggbarplot(auc_df_all_linear_iPSYCH2_vs_iPSYCH1, 
  x = "Features", 
  y = y_column, 
  fill = fill_column, 
  position = position_dodge(width = bar_width, preserve = "single"), 
  width = bar_width, 
  add = "", 
  facet.by = "Diagnosis",
  color = "black") +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    # subtitle = "Within-model p-value H1: performance on iPSYCH2 < performance on test_iPSYCH1",
    y = bquote(Delta * AUROC * .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
geom_errorbar(
    aes(
      ymin = !!sym(paste0("lower_", metric_name)),
      ymax = !!sym(paste0("upper_", metric_name)),
      group = interaction(!!sym(x_column), !!sym(fill_column))
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  scale_y_reverse(breaks = seq(y_min, y_max, by = 0.02)) +
  my_theme +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = features_colors_palette) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.background = element_blank(),
    plot.subtitle = element_text(size = 14)
    # legend.position = "none"
  ) +
  facet_wrap(~Diagnosis, nrow = 1) +
  geom_text(
    aes(
      x = Features, 
      y = upper_diff+0.01, 
      label = p.adj,
      group = Diagnosis
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5,
    size = pvalue_size_fig3
  );p_auc_perf_drop_all_linear_iPSYCH2

ggsave(file.path(figs_path, "auc_perf_drop_all_linear_iPSYCH2.pdf"), p_auc_perf_drop_all_linear_iPSYCH2 + theme(legend.position = "none"), width = 80, height = 30, units = "cm")

```

## R2 Linear perf drop iPSYCH2 vs test_iPSYCH1 all features
Here I want all the features but showing their drop in performance of their linear integration in iPSYCH2 vs test_iPSYCH1
```{r}

#-
features  <- c(
  "GLN",
  "bigstatsr",
  "GLN+bigstatsr",
  "PGS",
  "FGRS",
  "PGS+FGRS",
  "GLN+PGS+FGRS",
  "bigstatsr+PGS+FGRS",
  "GLN+bigstatsr+PGS+FGRS"
)
metrics_avg %>% count(Features)
# R2
# Calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the average values
r2_df_all_linear_iPSYCH2_vs_test_iPSYCH1 <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter_out_ADHD() %>% filter(Split_eval != "valid")%>% filter(Model_type == "Linear" | (Model_type == "DL" & Model_genotype == "GLN")) %>% filter(CV == "5-CV"),
  features_to_compare = features,
  cols_to_remove = c("CV", "Label", "Model_type", "Split_trained"),
  category_col = "Split_eval",
  metrics_cols = metrics_r2,
  split1 = "test_iPSYCH1",
  split2 = "iPSYCH2"
) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))); r2_df_all_linear_iPSYCH2_vs_test_iPSYCH1

r2_df_all_linear_iPSYCH2_vs_test_iPSYCH1 %>% count(Model, Features)


# Format the dataframe to calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the bootstrap replicates
column_replicates <- "r2_replicates"
r2_formatted_drop_all_linear_iPSYCH2  <- metrics_dl_vs_linear_repl %>%          
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(CV == "5-CV") %>%
          filter(Model_type == "Linear" | (Model_type == "DL" & Model_genotype == "GLN")) %>% 
          # filter(Features %in% runs_to_compare[[1]]) %>%
          # mutate(Features = fct_relevel(Features, runs_to_compare[[1]])) %>% 
          mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
          select(Diagnosis, Features, Split_eval, Split_trained, CV, all_of(column_replicates)) %>% 
          pivot_wider(names_from = Split_eval, values_from = all_of(column_replicates)) %>% 
          mutate(!!sym(column_replicates) := map2(iPSYCH2, test_iPSYCH1, ~ .x - .y))

# ----------------------------------
# Within-model pvalues
# ----------------------------------
# Get the real p-values for the GLN and bigstatsr models
r2_pvalues_all_linear_iPSYCH2_vs_iPSYCH1 <- r2_pvalues_same_feature_same_model_different_eval %>% 
filter_out_ADHD() %>% 
filter(Model_type == "Linear" | (Model_type == "DL" & Features == "GLN"))



#-
# Format the dataframe to be compatible with geom_text
r2_df_all_linear_iPSYCH2_vs_iPSYCH1  <- r2_df_all_linear_iPSYCH2_vs_test_iPSYCH1 %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
left_join(r2_pvalues_all_linear_iPSYCH2_vs_iPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(Features = fct_relevel(Features, features)) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

saveRDS(r2_df_all_linear_iPSYCH2_vs_iPSYCH1, file.path(final_folder, "r2_df_all_linear_iPSYCH2_vs_iPSYCH1.rds"))



#-
# Barplot
x_column <- "Diagnosis"
y_column <- "diff_avg"
fill_column <- "Features"
metric_name <- "diff"
bar_width  <- 0.9
y_min  <- -0.05
y_max  <- 0.12


p_r2_perf_drop_all_linear_iPSYCH2  <- ggbarplot(r2_df_all_linear_iPSYCH2_vs_iPSYCH1, 
  x = "Features", 
  y = y_column, 
  fill = fill_column, 
  position = position_dodge(width = bar_width, preserve = "single"), 
  width = bar_width, 
  add = "", 
  facet.by = "Diagnosis",
  color = "black") +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    # subtitle = "Within-model p-value H1: performance on iPSYCH2 < performance on test_iPSYCH1",
    y = bquote(Delta * R^2 * .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
geom_errorbar(
    aes(
      ymin = !!sym(paste0("lower_", metric_name)),
      ymax = !!sym(paste0("upper_", metric_name)),
      group = interaction(!!sym(x_column), !!sym(fill_column))
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  scale_y_reverse(breaks = seq(y_min, y_max, by = 0.02)) +
  my_theme +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = features_colors_palette) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.background = element_blank(),
    plot.subtitle = element_text(size = 14)
    # legend.position = "none"
  ) +
  facet_wrap(~Diagnosis, nrow = 1) +
  geom_text(
    aes(
      x = Features, 
      y = upper_diff+0.01, 
      label = p.adj,
      group = Diagnosis
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5,
    size = pvalue_size_fig3
  );p_r2_perf_drop_all_linear_iPSYCH2

ggsave(file.path(figs_path, "r2_perf_drop_all_linear_iPSYCH2.pdf"), p_r2_perf_drop_all_linear_iPSYCH2 + theme(legend.position = "none"), width = 80, height = 30, units = "cm")

```


## AUC DL perf drop iPSYCH2 vs test_iPSYCH1 all features
Here I want all the features but showing their drop in performance of their linear integration in iPSYCH2 vs test_iPSYCH1
```{r}


features  <- c(
  "GLN",
  "bigstatsr",
  "GLN+bigstatsr",
  "PGS",
  "FGRS",
  "PGS+FGRS",
  "GLN+PGS+FGRS",
  "bigstatsr+PGS+FGRS",
  "GLN+bigstatsr+PGS+FGRS"
)
metrics_avg %>% count(Features)
# AUROC
# Calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the average values
auc_drop_all_DL_iPSYCH2 <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter_out_ADHD() %>% filter(Model_type == "DL" | (Model_type == "Linear" & Model_genotype == "bigstatsr")) %>% filter(CV == "5-CV"),
  features_to_compare = features,
  cols_to_remove = c("CV", "Label", "Model_type", "Split_trained"),
  category_col = "Split_eval",
  metrics_cols = metrics_auroc,
  split1 = "test_iPSYCH1",
  split2 = "iPSYCH2"
) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))

auc_drop_all_DL_iPSYCH2 %>% count(Features, Model)

# metrics_dl_vs_linear_avg %>% filter_out_ADHD() %>% 
# # filter(Model_type == "DL" | (Model_type == "Linear" & Model_genotype == "bigstatsr" & Model != "EIR"))  %>% 
# filter(Model_type == "DL" | (Model_type == "Linear" & Model_genotype == "bigstatsr"))  %>% 
# filter(CV == "5-CV") %>% 
# filter(Features %in% c("GLN","GLN_on_geno","bigstatsr","bigstatsr_on_geno")) %>% 
# count(Model, Model_type, Features)

# Format the dataframe to calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the bootstrap replicates
column_replicates <- "auc_replicates"
auc_formatted_drop_all_DL_iPSYCH2  <- metrics_dl_vs_linear_repl %>%          
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(CV == "5-CV") %>%
          filter(Model_type == "DL") %>% 
          # filter(Features %in% runs_to_compare[[1]]) %>%
          # mutate(Features = fct_relevel(Features, runs_to_compare[[1]])) %>% 
          mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
          select(Diagnosis, Features, Split_eval, Split_trained, CV, all_of(column_replicates)) %>% 
          pivot_wider(names_from = Split_eval, values_from = all_of(column_replicates)) %>% 
          mutate(!!sym(column_replicates) := map2(iPSYCH2, test_iPSYCH1, ~ .x - .y))

# ----------------------------------
# Within-model pvalues
# ----------------------------------
# Get the real p-values for the GLN and bigstatsr models
auc_pvalues_all_DL_iPSYCH2_vs_iPSYCH1 <- auc_pvalues_same_feature_same_model_different_eval %>% 
filter_out_ADHD() %>% 
filter(Model_type == "DL" | (Model_type == "Linear" & Features == "bigstatsr"))

#-
# Format the dataframe to be compatible with geom_text
auc_df_all_DL_iPSYCH2_vs_iPSYCH1  <- auc_drop_all_DL_iPSYCH2 %>% 
# auc_drop_all_DL_iPSYCH2 %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
left_join(auc_pvalues_all_DL_iPSYCH2_vs_iPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(Features = fct_relevel(Features, features)) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

# create directory
saveRDS(auc_df_all_DL_iPSYCH2_vs_iPSYCH1, file.path(final_folder, "auc_df_all_DL_iPSYCH2_vs_iPSYCH1.rds"))

#-
# Barplot
x_column <- "Diagnosis"
y_column <- "diff_avg"
fill_column <- "Features"
metric_name <- "diff"
bar_width  <- 0.9
y_min  <- -0.05
y_max  <- 0.12


p_auc_perf_drop_all_DL_iPSYCH2  <- ggbarplot(auc_df_all_DL_iPSYCH2_vs_iPSYCH1, 
  x = "Features", 
  y = y_column, 
  fill = fill_column, 
  position = position_dodge(width = bar_width, preserve = "single"), 
  width = bar_width, 
  add = "", 
  facet.by = "Diagnosis",
  color = "black") +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    # subtitle = "Within-model p-value H1: performance on iPSYCH2 < performance on test_iPSYCH1",
    y = bquote(Delta * AUROC * .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
geom_errorbar(
    aes(
      ymin = !!sym(paste0("lower_", metric_name)),
      ymax = !!sym(paste0("upper_", metric_name)),
      group = interaction(!!sym(x_column), !!sym(fill_column))
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  scale_y_reverse(breaks = seq(y_min, y_max, by = 0.02)) +
  my_theme +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = features_colors_palette) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.background = element_blank(),
    plot.subtitle = element_text(size = 14)
    # legend.position = "none"
  ) +
  facet_wrap(~Diagnosis, nrow = 1) +
  geom_text(
    aes(
      x = Features, 
      y = upper_diff+0.01, 
      label = p.adj,
      group = Diagnosis
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5,
    size = pvalue_size_fig3
  );p_auc_perf_drop_all_DL_iPSYCH2

ggsave(file.path(figs_path, "auc_perf_drop_all_DL_iPSYCH2.pdf"), p_auc_perf_drop_all_DL_iPSYCH2 + theme(legend.position = "none"), width = 80, height = 30, units = "cm")

```

## R2 DL perf drop iPSYCH2 vs test_iPSYCH1 all features
Here I want all the features but showing their drop in performance of their linear integration in iPSYCH2 vs test_iPSYCH1
```{r}


features  <- c(
  "GLN",
  "bigstatsr",
  "GLN+bigstatsr",
  "PGS",
  "FGRS",
  "PGS+FGRS",
  "GLN+PGS+FGRS",
  "bigstatsr+PGS+FGRS",
  "GLN+bigstatsr+PGS+FGRS"
)
metrics_avg %>% count(Features)
# R2
# Calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the average values
r2_drop_all_DL_iPSYCH2 <- calculate_performance_diff(
  metrics_df = metrics_dl_vs_linear_avg %>% filter_out_ADHD() %>% filter(Model_type == "DL" | (Model_type == "Linear" & Model_genotype == "bigstatsr")) %>% filter(CV == "5-CV" & Split_eval != "valid"),
  features_to_compare = features,
  cols_to_remove = c("CV", "Label", "Model_type", "Split_trained"),
  category_col = "Split_eval",
  metrics_cols = metrics_r2,
  split1 = "test_iPSYCH1",
  split2 = "iPSYCH2"
) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis))))


# Format the dataframe to calculate the difference in performance between test_iPSYCH1 and iPSYCH2 for each model using the bootstrap replicates
column_replicates <- "r2_replicates"
r2_formatted_drop_all_DL_iPSYCH2  <- metrics_dl_vs_linear_repl %>%          
          filter_out_ADHD() %>%
          filter(Split_trained == "train_valid") %>%
          filter(CV == "5-CV") %>%
          filter(Model_type == "DL") %>% 
          # filter(Features %in% runs_to_compare[[1]]) %>%
          # mutate(Features = fct_relevel(Features, runs_to_compare[[1]])) %>% 
          mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
          select(Diagnosis, Features, Split_eval, Split_trained, CV, all_of(column_replicates)) %>% 
          pivot_wider(names_from = Split_eval, values_from = all_of(column_replicates)) %>% 
          mutate(!!sym(column_replicates) := map2(iPSYCH2, test_iPSYCH1, ~ .x - .y))

# ----------------------------------
# Within-model pvalues
# ----------------------------------
# Get the real p-values for the GLN and bigstatsr models
r2_pvalues_all_DL_iPSYCH2_vs_iPSYCH1 <- r2_pvalues_same_feature_same_model_different_eval %>% 
filter_out_ADHD() %>% 
filter(Model_type == "DL" | (Model_type == "Linear" & Features == "bigstatsr"))




#-
# Format the dataframe to be compatible with geom_text
r2_df_all_DL_iPSYCH2_vs_iPSYCH1  <- r2_drop_all_DL_iPSYCH2 %>% 
mutate(Features = str_replace_all(Features, "_on_geno", "")) %>% 
left_join(r2_pvalues_all_DL_iPSYCH2_vs_iPSYCH1, by = c("Diagnosis", "Features")) %>% 
mutate(Diagnosis = factor(Diagnosis, levels = sort(unique(Diagnosis)))) %>% 
mutate(Features = fct_relevel(Features, features)) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = FALSE), "ns"))

saveRDS(r2_df_all_DL_iPSYCH2_vs_iPSYCH1, file.path(final_folder, "r2_df_all_DL_iPSYCH2_vs_iPSYCH1.rds"))

#-
# Barplot
x_column <- "Diagnosis"
y_column <- "diff_avg"
fill_column <- "Features"
metric_name <- "diff"
bar_width  <- 0.9
y_min  <- -0.05
y_max  <- 0.12


p_r2_perf_drop_all_DL_iPSYCH2  <- ggbarplot(r2_df_all_DL_iPSYCH2_vs_iPSYCH1, 
  x = "Features", 
  y = y_column, 
  fill = fill_column, 
  position = position_dodge(width = bar_width, preserve = "single"), 
  width = bar_width, 
  add = "", 
  facet.by = "Diagnosis",
  color = "black") +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    # subtitle = "Within-model p-value H1: performance on iPSYCH2 < performance on test_iPSYCH1",
    y = bquote(Delta * R^2 * .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
geom_errorbar(
    aes(
      ymin = !!sym(paste0("lower_", metric_name)),
      ymax = !!sym(paste0("upper_", metric_name)),
      group = interaction(!!sym(x_column), !!sym(fill_column))
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  scale_y_reverse(breaks = seq(y_min, y_max, by = 0.02)) +
  my_theme +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = features_colors_palette) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.background = element_blank(),
    plot.subtitle = element_text(size = 14)
    # legend.position = "none"
  ) +
  facet_wrap(~Diagnosis, nrow = 1) +
  geom_text(
    aes(
      x = Features, 
      y = upper_diff+0.01, 
      label = p.adj,
      group = Diagnosis
    ),
    position = position_dodge(width = bar_width),
    vjust = -0.5,
    size = pvalue_size_fig3
  );p_r2_perf_drop_all_DL_iPSYCH2

ggsave(file.path(figs_path, "r2_perf_drop_all_DL_iPSYCH2.pdf"), p_r2_perf_drop_all_DL_iPSYCH2 + theme(legend.position = "none"), width = 80, height = 30, units = "cm")

```

## AUROC integrations Linear vs DL drop perf iPSYCH2 vs test_iPSYCH1
```{r}

auc_df_all_linear_iPSYCH2_vs_iPSYCH1 <- readRDS(file.path(final_folder, "auc_df_all_linear_iPSYCH2_vs_iPSYCH1.rds"))
auc_df_all_DL_iPSYCH2_vs_iPSYCH1 <- readRDS(file.path(final_folder, "auc_df_all_DL_iPSYCH2_vs_iPSYCH1.rds"))
#-
auc_df_all_DL_linear_iPSYCH2_vs_iPSYCH1 <- bind_rows(auc_df_all_linear_iPSYCH2_vs_iPSYCH1, auc_df_all_DL_iPSYCH2_vs_iPSYCH1) %>% 
filter(str_detect(Features, "GLN") & str_detect(Features, "\\+")) %>% 
mutate(Integration = fct_relevel(Integration, c("Linear", "DL"))) %>% 
select(-c(contains("p_value_bonf"), contains("p_value_fdr"))) %>% 
mutate(p_value_fdr_all = p.adjust(p_value, method = "fdr")) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = TRUE), "ns"))



#-
bar_width  <- 0.7
pvalue_size_fig3 <- 7


auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear  <- readRDS(file.path(final_folder, "auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
metrics_linear_test  <- readRDS(file.path(final_folder, "metrics_linear_test.rds"))


p_auc_tabs_linear_test  <- plot_barplots_ggpubr(dataframe = metrics_linear_test,
  x_column = "Features",
  fill_column = "Features",
  fill_label = "Features",
  title = "",
  x_title = "",
  metrics = metrics_auroc,
  y_title = "AUROC",
  # metrics = metrics_auroc,
  # y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Diagnosis",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right", # Remove legend by setting position to "none"
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  # pvalue_size = 9,
  fill_colors = features_colors_palette,
  # padding = -0.001
  padding = 0,
  bar_width = 0.9,
  theme = my_theme
  # expand = c(0, 0)
) + 
# add horizontal dashed line at 0.5
geom_hline(yintercept = 0.5, linetype = "dashed", color = "black")+
# theme(aspect.ratio = 1)+
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.5)) +
# theme(axis.ticks.x = element_line()) +
theme(strip.background = element_blank()) +
# rotate 90 degrees
theme(axis.text.x = element_text(angle = 90, hjust = 1))+
scale_y_continuous(breaks = seq(0.5, 0.77, by = 0.1), limits = c(0, 0.77), expand = c(0, 0))+
coord_cartesian(ylim = c(0.495, 0.74))+
scale_y_continuous(breaks = c(seq(0.5, 1, by = 0.05)))+
stat_pvalue_manual(
auc_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear %>% 
mutate(y.position = 
case_when(
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr" ~ 0.65,
  Diagnosis == "ADHD" & group2 == "PGS+FGRS" ~ 0.66,
  Diagnosis == "ADHD" & group2 == "GLN+PGS+FGRS" ~ 0.705,
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.72,

  Diagnosis == "ASD" & group2 == "GLN+bigstatsr" ~ 0.59,
  Diagnosis == "ASD" & group2 == "PGS+FGRS" ~ 0.61,
  Diagnosis == "ASD" & group2 == "GLN+PGS+FGRS" ~ 0.63,
  Diagnosis == "ASD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.645,

  Diagnosis == "BIP" & group2 == "GLN+bigstatsr" ~ 0.60,
  Diagnosis == "BIP" & group2 == "PGS+FGRS" ~ 0.71,
  Diagnosis == "BIP" & group2 == "GLN+PGS+FGRS" ~ 0.72,
  Diagnosis == "BIP" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.73,

  Diagnosis == "MDD" & group2 == "GLN+bigstatsr" ~ 0.61,
  Diagnosis == "MDD" & group2 == "PGS+FGRS" ~ 0.66,
  Diagnosis == "MDD" & group2 == "GLN+PGS+FGRS" ~ 0.67,
  Diagnosis == "MDD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.68,

  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr" ~ 0.62,
  Diagnosis == "SCZ" & group2 == "PGS+FGRS" ~ 0.675,
  Diagnosis == "SCZ" & group2 == "GLN+PGS+FGRS" ~ 0.68,
  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.69,
  TRUE ~ y.position
)
),
label = "p_fdr", 
tip.length = 0.01, 
size = pvalue_size_fig3)+
# remove legend
theme(legend.position = "none")



perf_drop_test_auc_tabs <- readRDS(file.path(final_folder, "perf_drop_test_auc_tabs.rds"))

bar_width  <- 0.7
pvalue_size_fig3 <- 7
p_perf_drop_tabs_test_DL_linear_auc <- ggbarplot(perf_drop_test_auc_tabs,
  x = "Features",
  y = "diff_avg",
  fill = "Features",
  position = position_dodge(width = bar_width, preserve = "single"),
  width = bar_width,
  add = "",
  color = "black"
) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * AUROC ~ .(suffix_delta_model)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Features
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.006,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = +0.7, # Adjust vertical spacing
    size = pvalue_size_fig3
  ) +
  scale_y_reverse() +
  my_theme +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  facet_grid(~Diagnosis, scales = "free_x")


p_auc_perf_drop_tabs_iPSYCH2_linear_vs_DL_fig3C  <- ggplot(auc_df_all_DL_linear_iPSYCH2_vs_iPSYCH1,
  aes(
    x = Features,
    y = diff_avg,
    fill = Features,
    pattern = Integration  # Add pattern aesthetic
  )
) +
# geom_bar(
#     position = position_dodge(width = bar_width, preserve = "single"),
#     width = bar_width,
#     color = "black",
#     stat = "identity"
#   ) +
geom_bar_pattern(
    position = position_dodge(width = bar_width, preserve = "single"),
    width = bar_width,
    color = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    stat = "identity",
    pattern_key_scale_factor = 0.5
  ) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * AUROC ~ .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Integration
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.006,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = +0.7, # Adjust vertical spacing
    size = pvalue_size_fig3
  ) +
  scale_y_reverse() +
  my_theme +
  guides(fill = "none", pattern = guide_legend(title = "Integration", override.aes = list(fill = features_colors_palette["GLN+PGS+FGRS"]))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  # Using different patterns that are better supported in ggpattern
  scale_pattern_manual(values = c("Linear" = "stripe", "DL" = "crosshatch")) +
  facet_grid(~Diagnosis, scales = "free_x")+
  theme(legend.position = "bottom")




row2 <- ((p_perf_drop_tabs_test_DL_linear_auc + theme(legend.position = "none")) | p_auc_perf_drop_tabs_iPSYCH2_linear_vs_DL_fig3C) + plot_layout(widths = c(1, 2))



fig3_auc <- 
  (p_auc_tabs_linear_test + theme(legend.position = "none")) / 
  row2 +
  plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(face = "bold"))

# ggsave(glue("plots_paper_final/figs_{date}/perf_drop_tabs_test_DL_linear_auc.pdf"), p_perf_drop_tabs_test_DL_linear_auc, width = 50, height = 30, units = "cm")

width  <- 90
height <- 60
ggsave(file.path("/Users/lmz306", "fig3_auc.pdf"), fig3_auc, width = width, height = height, units = "cm")

# cairo_pdf(glue("plots_paper_final/figs_{date}/fig3_auc_{width}x{height}.pdf"), width = width / 2.54, height = height / 2.54)
# print(fig3_auc)
# dev.off()




```

## R2 integrations Linear vs DL drop perf iPSYCH2 vs test_iPSYCH1

```{r}

#-

r2_df_all_linear_iPSYCH2_vs_iPSYCH1  <- readRDS(file.path(final_folder, "r2_df_all_linear_iPSYCH2_vs_iPSYCH1.rds"))

r2_df_all_DL_iPSYCH2_vs_iPSYCH1  <- readRDS(file.path(final_folder, "r2_df_all_DL_iPSYCH2_vs_iPSYCH1.rds"))

r2_df_all_DL_linear_iPSYCH2_vs_iPSYCH1 <- bind_rows(r2_df_all_linear_iPSYCH2_vs_iPSYCH1, r2_df_all_DL_iPSYCH2_vs_iPSYCH1) %>% 
filter(str_detect(Features, "GLN") & str_detect(Features, "\\+")) %>% 
mutate(Integration = fct_relevel(Integration, c("Linear", "DL"))) %>% 
select(-c(contains("p_value_bonf"), contains("p_value_fdr"))) %>% 
mutate(p_value_fdr_all = p.adjust(p_value, method = "fdr")) %>% 
mutate(p.adj = ifelse(p_value_fdr_all < 0.05, format(p_value_fdr_all, digits = 2, scientific = TRUE), "ns"))

bar_width  <- 0.7



r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear  <- readRDS(file.path(final_folder, "r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear.rds"))
metrics_linear_test  <- readRDS(file.path(final_folder, "metrics_linear_test.rds"))

p_r2_tabs_linear_test  <- plot_barplots_ggpubr(dataframe = metrics_linear_test,
  x_column = "Features",
  fill_column = "Features",
  fill_label = "Features",
  title = "",
  x_title = "",
  metrics = metrics_r2,
  y_title = "R2",
  # metrics = metrics_auroc,
  # y_title = "AUROC",
  facet = TRUE,
  facet_type = "grid",
  facet_var = "Diagnosis",
  x_angle = 0,
  x_hjust = 1,
  legend_position = "right",
  text_size = 11,
  title_hjust = 0.5,
  add_text = FALSE,
  fill_colors = features_colors_palette,
  # padding = -0.001
  padding = 0,
  bar_width = 0.9,
  theme = my_theme
  # expand = c(0, 0)
) + 
# add horizontal dashed line at 0.5
# theme(aspect.ratio = 1)+
theme(axis.text.x = element_text(hjust = 0.5, vjust = 0.5)) +
theme(strip.background = element_blank()) +
theme(axis.text.x = element_text(angle = 90, hjust = 1))+
stat_pvalue_manual(
r2_pvalues_formatted_nested_models_test_iPSYCH1_toplot_linear %>% 
mutate(y.position = 
case_when(
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr" ~ 0.06,
  Diagnosis == "ADHD" & group2 == "PGS+FGRS" ~ 0.09,
  Diagnosis == "ADHD" & group2 == "GLN+PGS+FGRS" ~ 0.105,
  Diagnosis == "ADHD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.115,

  Diagnosis == "ASD" & group2 == "GLN+bigstatsr" ~ 0.02,
  Diagnosis == "ASD" & group2 == "PGS+FGRS" ~ 0.04,
  Diagnosis == "ASD" & group2 == "GLN+PGS+FGRS" ~ 0.05,
  Diagnosis == "ASD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.06,

  Diagnosis == "BIP" & group2 == "GLN+bigstatsr" ~ 0.02,
  Diagnosis == "BIP" & group2 == "PGS+FGRS" ~ 0.225,
  Diagnosis == "BIP" & group2 == "GLN+PGS+FGRS" ~ 0.235 ,
  Diagnosis == "BIP" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.245,

  Diagnosis == "MDD" & group2 == "GLN+bigstatsr" ~ 0.03,
  Diagnosis == "MDD" & group2 == "PGS+FGRS" ~ 0.095,
  Diagnosis == "MDD" & group2 == "GLN+PGS+FGRS" ~ 0.105,
  Diagnosis == "MDD" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.115,

  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr" ~ 0.025,
  Diagnosis == "SCZ" & group2 == "PGS+FGRS" ~ 0.09,
  Diagnosis == "SCZ" & group2 == "GLN+PGS+FGRS" ~ 0.105,
  Diagnosis == "SCZ" & group2 == "GLN+bigstatsr+PGS+FGRS" ~ 0.115,
  TRUE ~ y.position
)
),
label = "p_fdr", 
tip.length = 0.01, 
size = pvalue_size_fig3)

perf_drop_test_r2_tabs <- readRDS(file.path(final_folder, "perf_drop_test_r2_tabs.rds"))
bar_width  <- 0.7
p_perf_drop_tabs_test_DL_linear_r2 <- ggbarplot(perf_drop_test_r2_tabs,
  x = "Features",
  y = "diff_avg",
  fill = "Features",
  position = position_dodge(width = bar_width, preserve = "single"),
  width = bar_width,
  add = "",
  color = "black"
) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * R^2 * .(suffix_delta_model)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Features
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.006,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = +0.7, # Adjust vertical spacing
    size = pvalue_size_fig3 
  ) +
  scale_y_reverse() +
  my_theme +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  facet_grid(~Diagnosis, scales = "free_x")

p_r2_perf_drop_tabs_iPSYCH2_linear_vs_DL_fig3C  <- ggplot(r2_df_all_DL_linear_iPSYCH2_vs_iPSYCH1,
  aes(
    x = Features,
    y = diff_avg,
    fill = Features,
    pattern = Integration  # Add pattern aesthetic
  )
) +
geom_bar_pattern(
    position = position_dodge(width = bar_width, preserve = "single"),
    width = bar_width,
    color = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    stat = "identity",
    pattern_key_scale_factor=0.5
  ) +
  geom_hline(yintercept = 0) +
  labs(
    title = "",
    y = bquote(Delta * R^2 * .(suffix_delta_eval)),
    x = "",
    fill = "Features"
  ) +
  geom_errorbar(
    aes(
      ymin = lower_diff,
      ymax = upper_diff,
      group = Integration
    ),
    width = 0.2,
    position = position_dodge(width = bar_width)
  ) +
  geom_text(
    aes(
      x = Features,
      y = upper_diff+0.006,
      label = p.adj
    ),
    position = position_dodge(width = bar_width),
    vjust = +0.7, # Adjust vertical spacing
    size = pvalue_size_fig3
  ) +
  scale_y_reverse() +
  my_theme +
  guides(fill = "none", pattern = guide_legend(title = "Integration", override.aes = list(fill = features_colors_palette["GLN+PGS+FGRS"]))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_blank()) +
  scale_fill_manual(values = features_colors_palette) +
  scale_pattern_manual(values = c("Linear" = "stripe", "DL" = "crosshatch")) +
  facet_grid(~Diagnosis, scales = "free_x")+
    theme(legend.position = "bottom")


row2  <- ((p_perf_drop_tabs_test_DL_linear_r2 + theme(legend.position = "none")) | p_r2_perf_drop_tabs_iPSYCH2_linear_vs_DL_fig3C) + plot_layout(widths = c(1, 2))

p_tabs_test_r2  <- (p_r2_tabs_linear_test + theme(legend.position = "none")) / row2 + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(face = "bold"))

# ggsave(glue("plots_paper_final/figs_{date}/perf_drop_tabs_test_DL_linear_r2.pdf"), p_perf_drop_tabs_test_DL_linear_r2, width = 50, height = 30, units = "cm")

width  <- 90
height <- 60
ggsave(file.path("/Users/lmz306", "FigS5_r2.pdf"), p_tabs_test_r2, width = width, height = height, units = "cm")



```
